### 2017-08-04 Manuscript.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{\normalsize{Biometry of the invasive lionfish (\textit{Pterois volitans}) in the Central Mexican Caribbean, and a review of allometric growth parameters across the invasion range}}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers

\begin{document}
\maketitle

Juan Carlos Villaseñor-Derbez\textsuperscript{1}

\textsuperscript{1} Bren School of Environmental Sciences and
Management, University of California Santa Barbara, Santa Barbara,
California, U.S.

\textbf{Full postal address}\\
Bren Hall\\
University of California at Santa Barbara\\
Santa Barbara\\
CA, 93106

\textbf{Corresponding author information:}\\
Juan Carlos Villaseñor-Derbez\\
Bren Hall\\
University of California at Santa Barbara\\
Santa Barbara\\
CA, 93106
\href{mailto:jvillasenor@bren.ucsb.edu}{\nolinkurl{jvillasenor@bren.ucsb.edu}}
+1 207 205 8435

\textbf{Running title}\\
Allometric growth of Pterois volitans

\clearpage

\textbf{Title of contribution}: Biometry of the invasive lionfish
(\emph{Pterois volitans}) in the Central Mexican Caribbean, and a review
of allometric growth parameters across the invasion range

\textbf{Abstract}\\
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
North-Western Atlantic and the Caribbean, likely introduced through
liberation of aquarium-kept organisms. Due to its threat to local
biodiversity, the speed of their spread, and its difficulty of
management, their presence in these waters has been labeled as a major
marine invasion. In order to better manage the invasion, inform lionfish
removal programs, predict their possible impacts, and estimate biomass
available for harvest, we must be able to accurately estimate their
total biomass. This study reports allometric growth parameters
(\(a = 3.2056\times 10^{-6}; b = 3.235\)) for organisms sampled in the
central portion of the Mexican Caribbean. It also identifies that, for a
given length, organisms from the Caribbean tend to have lower weights
than those from the Gulf of Mexicoand Atlantic, indicating that
weight-at-length is spacialy variable. In addition to the new pair of
parameters provided in this study, other 13 pairs of allometric growth
parameters were used to calculate expected biomass from total length
observations. Mean expected-to-observed biomass ratios by study ranged
between 0.98 and 1.75, with overall values ranging from 0.44 to 3.51.
The large differences between observed and expected biomass when using
parameters from other locations highlights the importance of using
site-specific parameters to estimate biomass from length observations.
This becomes increasingly important when estimating biomass available
for harvest or effectiveness of lionfish removal efforts. Additionally,
guidelines on reporting allometric parameters are provided as a
suggestion for future research are provided.

\textbf{Key words}\\
lionfish, Pterois volitans, length-weight relationship, allometric
growth, Mexico

\textbf{Resumen}\\
Los peces león (\emph{Pterois volitans/miles}) son una especiei nvasora
en el Atlántico Noroeste y el Caribe, probablemente introducidos por
medio de la liberación de organismos de ornato. Dada la amenaza que
estos representan para la biodiversidad local, la velocidad de su
invasión y dificultad de manejo, su presencia ha sido catalogada como
una importante invasión marina. Para manejar mejor la invasión, proveer
información a los programas de remoción, predecir los posibles impactos
y estimar biomasa disponible para la extracción, debemos ser capaces de
estimar su biomasa con precisión. Este estudio presenta parámetros de
crecimiento alométrico (\(a = 3.2056\times 10^{-6}; b = 3.235\)) para
organismos muestreados en la porción central del Caribe Mexicano.
También identifica que para una misma longitud, los organismos del
Caribe tienden a ser menos pesados que aquellos del Golfo de México y
Atlántico, indicando variaciones espaciales en la relación
peso-longitud. Además de los parámetros descritos en este estudio, se
utilizaron otros 13 pares de parámetros de crecimiento alométrico para
estimar biomasa a partir de observaciones de talla. La razón entre
biomasa observada y biomasa esperada varió, en promedio por estudio,
entre 0.98 y 1.75, con valores extremos de 0.44 y 3.51. Las grandes
diferencias entre biomasa observada y esperada al usar parámetros de
otras regiones resalta la importancia de utilizar parámetros específicos
a la zona para estimar biomasa a partir de observaciones de talla.
Adicionalmente, se proveen sugerencias para estandarizar la manera en
que los parámetros de crecimiento alométrico se reportan.

\textbf{Palabras clave} pez león, Pterois volitans, relación
longitud-peso, crecimiento alométrico, México

\clearpage

\section{Introduction}\label{introduction}

At least 84\% of the marine eco-regions have reported the presence of an
invasive species (Molnar et al., 2008). These represent a major threat
to local biodiversity and the economic activities that depend on it,
like tourism or fisheries (Bax et al., 2003). Invasive species may also
threaten native species through competition (DAVIS, 2003) or predation.
By 2005, the economic cost of invasive species to the United States was
estimated at \$120 billion per year and nearly 42\% of species that have
been included in the Endangered or Threatened species list have been
labeled as such due to presence of invasive species (Pimentel, Zuniga \&
Morrison, 2005). This highlights the importance of understanding,
managing, and preventing ecological invasions.

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the North-Western Atlantic and the Caribbean, likely introduced
through liberation of aquarium-kept organisms (Betancur-R et al., 2011).
They are the first marine vertebrates to establish in North Atlantic
(Schofield, 2009, 2010) and Caribbean coasts (Sabido-Itza et al., 2016).
Lionfish have been widely reported in coral reefs (Aguilar-Perera \&
Tuz-Sulub, 2010), but also in other habitats such as estuaries (Jud et
al., 2011), mangroves (Barbour et al., 2010), areas with hard-bottoms
(Muñoz, Currin \& Whitfield, 2011), and mesophotic reefs (Andradi-Brown
et al., 2017). Due to its threat to local biodiversity, the speed of
their spread, and its difficulty of management, their presence in these
waters has been labeled as a major marine invasion (Hixon et al., 2016).

A significant amount of research has been done to describe lionfish
feeding ecology in North Carolina (Muñoz, Currin \& Whitfield, 2011),
the Bahamas (Morris \& Akins, 2009; Cote et al., 2013), Northern Gulf of
Mexico (Dahl \& Patterson, 2014), Mexican Caribbean (Valdez-Moreno et
al., 2012; Villaseñor-Derbez \& Herrera-P\a'erez, 2014), Belize
(Hackerott et al., 2017), and Costa Rica (Sandel et al., 2015). Their
feeding behavior and high consumption rates can reduce recruitment
(Albins \& Hixon, 2008) and population sizes (Green et al., 2012) of
native reef-fish species, and further the endangerment of critically
endangered reef fish (Rocha et al., 2015). (However, see Hackerott et
al. (2017) for a case where there was no evidence that lionfish affected
the density, richness, or composition of prey fishes). Major efforts
have also been made to understand the possible impacts of the invasion
by keeping track of its range through time (Schofield, 2009, 2010) and
predicting invasion ranges under climate change scenarios (Grieve,
Curchitser \& Rykaczewski, 2016). By combining information from these
disciplines, researchers have been able to predict the trophic impacts
of lionfish (Arias-Gonzalez et al., 2011), which can then be translated
into ecosystem-level and economic impacts.

Seeking to reduce lionfish densities, governments and non-profit
organizations have promoted removal programs and incentivized its
consumption (Chin, Aiken \& Buddo, 2016). In some cases, these have
shown to significantly reduce -but not quite eliminate- lionfish
abundances at local scales (Sandel et al., 2015, Chin, Aiken \& Buddo
(2016), de Leon et al. (2013)). The rapid recovery rates exhibited by
lionfish (Barbour et al., 2011) and the persistent populations in
mesophotic coral ecosystems (Andradi-Brown et al., 2017) -which can
contribute with recruitment to shallow-water populations- make of
complete eradication through fishing effort an unlikely solution.
However, further incentivizing its consumption might create a demand big
enough to promote and sustain a stable fishery (Chin, Aiken \& Buddo,
2016), which can reduce local abundances and control -not eradicate- the
invasion while providing alternative livelihoods.

The feasibility of lionfish removal programs has been extensively
evaluated through field observations(Sandel et al., 2015, Chin, Aiken \&
Buddo (2016), de Leon et al. (2013); Usseglio et al., 2017) and
empirical modeling (Barbour et al., 2011; Morris, Shertzer \& Rice,
2011; Johnston \& Purkis, 2015). The latter measure changes in biomass
or density (Barbour et al., 2011; Johnston \& Purkis, 2015) in response
to increased mortality (\emph{i.e.} lionfish removal). In this case,
biomass represents the sum of all fish's individual weight. Total Weight
(TW) can be estimated from Total Length (TL) observations using the
allometric growth equation (\(TW = aTL^b\)). Parameters \(a\) and \(b\)
for this equation exist for North Carolina (Barbour et al., 2011),
Northern (Fogg et al., 2013) and Southern Gulf of Mexico (Aguilar-Perera
\& Quijano-Puerto, 2016), the Southern Mexican Caribbean (Sabido-Itza et
al., 2016), Little Cayman (Edwards, Frazer \& Jacoby, 2014), Jamaica
(Chin, Aiken \& Buddo, 2016), Bonaire (de Leon et al., 2013) and Costa
Rica (Sandel et al., 2015), but remain unavailable for the central
Mexican Caribbean. The weight-at-length of a species can vary across
regions as a response to biotic (\emph{e.g.} local food availability)
and abiotic (\emph{e.g.} water temperature) conditions (Johnson \&
Swenarton, 2016). Thus, when using biomass-informed models or estimating
biomass from length observations, it is important to use site-specific
parameters to obtain an accurate estimate. This is especially important
when research involves identifying the total biomass available for
harvest by fishers (Chin, Aiken \& Buddo, 2016) or the efficacy of
lionfish removals (Barbour et al., 2011; Morris, Shertzer \& Rice, 2011;
Johnston \& Purkis, 2015).

Here, I provide the first allometric growth parameters for the invasive
lionfish in the central Mexican Caribbean region. At the same time, I
highlight the importance of using site-specific parameters by estimating
biomass with parameters from other regions across the invasion range and
comparing them to observed biomass. I also provide other 13 standardized
parameters from eight studies through the invasion range, making them
readily accessible for future research. Finally, I discuss the way in
which allometric parameters are reported, and call for standardization
to facilitate their use.

\section{Materials and Methods}\label{materials-and-methods}

\subsection{Area of study}\label{area-of-study}

The study took place off the coasts of Playa del Carmen, in the Mexican
Caribbean (\textbf{Fig. 1}). The region represents the northernmost
section of the Mesoamerican Barrier Reef System (Ruiz-Zarate \&
Arias-Gonzalez, 2004). Coral reefs and mangroves are locally important
habitats that represent important sources of income in terms of
extractive (\emph{e.g.} recreational fishing) and non-extractive
(\emph{e.g.} SCUBA diving) activities related to tourism, the main
source of income to the local economy (Murray, 2007).

In the region, the reef profile has been described by Arias-Gonzalez
(1998), indicating that the reef lagoon extends about 500 m from the
coast, until the reef crest is reached. The reef becomes deeper, leading
to the reef front often found at 700 m from the coastline end extends
for an additional 300 m. At approximately 1000 m away from shore and 30
- 40 m depth, the reef leads to a drop-off. Along a perpendicular
profile to the coast, bands of reef are interrupted by sand patches at 8
- 12 m deep and 16-18 m deep. Along the coast, these reefs have been
reported to be under significant anthropogenic pressure, likely causing
a shift in structure and function (Bozec et al., 2008).

\subsection{Fish sampling}\label{fish-sampling}

Organisms sampled by Villaseñor-Derbez \& Herrera-P\a'erez (2014) for
stomach content analysis between May and August of 2010 were used in
this work. A total of 33 SCUBA immersions were performed in 10 sampling
sites along the coast (Fig. 1, Table 1). Sampling locations included
wall and carpet reefs at depths between 5.7 m and 38.1 m. All observed
organisms (n = 109) were collected using hand nets and numbered
collection bottles. The use of hand nets prevents any weight loss due to
bleeding(when collected with spears), and increases the range of sizes
that can be sampled, providing a representative sample of the
population. Information on depth and other comments were recorded in an
underwater slide. Depth was recorded by dive gauges held by divers as
safety procedures during the collections. Samples were frozen within 30
minutes of completing the dive and stored for posterior analysis in the
lab. For every organism Total Length (TL; mm) and Total Weight (TW; gr)
were recorded in the lab.

\subsection{Data analysis}\label{data-analysis}

The weight at length relationship between the observed variables is
described by the allometric growth function:

\begin{equation}TW = aTL^b\end{equation}

Where \(TW\) is the Total Weight (gr), \(TL\) is the observed Total
Length (mm), \(a\) is the ponderal index and \(b\) is the scaling
exponent or allometric parameter. When \(b = 3\), it is said that the
organism exhibits a perfect isometric growth. The dependent and
independent variables were transformed via base-10 logarithms so that
the equation is then:

\begin{equation}log_{10}(TW) = b\times log_{10}(TL) + log{10}(a)\end{equation}

To simplify this equation, we can re-write it as:

\begin{equation}Y = mX + c\end{equation}

Where \(Y = log_{10}(TW)\), \(X = log_{10}(TL)\), \(m = b\), and
\(c = log_{10}(a)\). Since \(b == m\), we will only use \(b\) throughout
the paper for simplicity. The coefficients (\(c\) and \(b\)) were
estimated with an Ordinary Least Square Regression and
heteroskedastic-robust standard errors. Both coefficients were tested
against the null hypothesis of no change (\emph{i.e.} \(H_0: c = 0\) and
\(H_0: b = 0\)). Additionally, the allometric parameter was tested
against the null hypothesis of isometric growth (\(H_0: b = 3\)).
Coefficients were tested with a two-tailed Student's t-test. The
significance of the regression was corroborated with an F-test.

Other allometric growth parameters were obtained from peer-reviewed
literature. Parameters were obtained forNorth Carolina (Barbour et al.,
2011), Northern (Fogg et al., 2013) and Southern Gulf of Mexico
(Aguilar-Perera \& Quijano-Puerto, 2016), the Southern Mexican Caribbean
(Sabido-Itza et al., 2016), Little Cayman (Edwards, Frazer \& Jacoby,
2014), Jamaica (Chin, Aiken \& Buddo, 2016), Bonaire (de Leon et al.,
2013) and Costa Rica (Sandel et al., 2015). Additionally, parameters
from Fishbase (Froese \& Pauly, 2016) were also included in this study.
When available, information on sampling methods, gender differentiation,
location, and depth ranges of each study was retrieved. Whenever gender
was not specified, it was assumed that the results were presented for
both genders pooled together. During the review process, it was noticed
that some papers indistinctly use \(a\) to report either the ponderal
index in eq. 1 or the y-intercept (\(c\)) in eq. 3, which might
sometimes be overlooked. Furthermore, some studies report their
parameters as mm-to-gr conversions, but a rapid evaluation of such
parameters indicates that they were estimated as cm-to-gr conversions.
Here, all parameters are reported as TL(mm) to TW(gr) conversions. When
required, values from other studies are transformed for consistency.

Since standard errors or confidence intervals were not provided for many
studies, it was impossible to test for statistical differences across
studies. Thus expected biomass for each organism was calculated with the
growth parameters estimated in this study, as well as those retrieved
from additional literature. All expected biomass values were divided by
the observed biomass to obtain a ratio that allowed rapid identification
of over- or underestimation with respect to the observed biomass.
Therefore, values lower than 1 indicate that the parameters
underestimate biomass, and values larger than 1 indicate overestimation.
Median and mean expected to observed biomass ratios were calculated for
each study.

All hypothesis testing was performed with a confidence level of
\(\alpha = 0.01\) in R version 3.4.0 (R Core Team, 2017). Data wrangling
was performed with the tidyverse package (Wickham, 2017). Maps were
created with a mix of functions from the sp (Pebesma \& Bivand, 2005),
rgdal (Bivand, Keitt \& Rowlingson, 2017), tmap (Tennekes, 2017a), and
tmaptools (Tennekes, 2017b) packages. Calculation of
heteroskedastic-robust standard errors was done with the sandwich
(Zeileis, 2004) and lmtest (Zeileis \& Hothorn, 2002) packages. Models
were manipulated with with the broom package (Robinson, 2017). Stargazer
(Hlavac, 2015), knitr (Xie, 2017), and kableExtra (Zhu, 2017) were used
to produce the tables, and RefManageR (McLean, 2014) was used to keep
track of citations. The manuscript was written in r markdown (Allaire et
al., 2017) and processed with the knitr package (Xie, 2017). Raw data
and code used in this work is available at
\url{github.com/jcvdav/lionfish_biometry}.

\section{Results}\label{results}

Organism TL ranged between 34 and 310 mm and TW between 0.3 and 397.7
gr. The smallest organism (TL = 34.00 mm) was also the lightest organism
(TW = 0.30 gr). However, the largest organism (TL = 310.00 mm) was not
the heaviest (TW = 303.70 gr), and the heaviest organism (TW = 397.70
gr) was 292.00 mm in total length. Kernell density plots (Fig. 2) show
the distribution for TL and TW. Both measures were positively skewed,
with skewness of 0.85 for TL and 2.19 for TW.

\subsection{Length-weight
relationship}\label{length-weight-relationship}

The model adjusted to eq. 3 estimated the coefficient values at
\(b = 3.2347391\) and \(c = -5.4940866\). Thus, TW (gr) can be
calculated from TL (mm) as a linear equation:
\(log_{10}(TW) = 3.2347391\times log_{10}(TL) -5.4940866\) or its
exponential form: \(TW = 3.2056297\times 10^{-6}\times TL^{3.2347391}\).
The intercept (\(c\)) and slope \((b)\) were significantly different
from zero (\(t(107) = -66.17; p<0.01\) and \(t(107) = 83.24; p<0.01\),
respectively), rejecting the null hypothesis of no change. Additionally,
the allometric factor (\(m = b\)) was significantly different from the
value of isometric growth of \(b = 3\) (\(t(107) = 6.04; p<0.01\)),
indicating that lionfish present allometric growth. More information on
model fit and confidence intervals for the estimated coefficients is
presented in Table 2. The relationship between Total Length and Total
Weight is presented in Figure 3.

\subsection{Comparison of allometric
parameters}\label{comparison-of-allometric-parameters}

From the eight peer-reviewed studies including information on growth
parameters for \emph{P. volitans} and Fishbase (Froese \& Pauly, 2016),
13 parameters were identified (Table 3). Two studies (Fogg et al., 2013;
Aguilar-Perera \& Quijano-Puerto, 2016) reported gender-level and pooled
parameters, while the rest of the studies always presented pooled
results. The smallest coefficient of determination was presented by
Chin, Aiken \& Buddo (2016) with \(R^2 = 0.8715\), while Sabido-Itza et
al. (2016) reported the highest value at \(R^2 = 0.9907\). These studies
presented information for organisms obtained at depths between 0.5 and
57 m. Two studies (Aguilar-Perera \& Quijano-Puerto, 2016; Chin, Aiken
\& Buddo, 2016) explicitly stated that their organisms were sampled with
pole spears. Five studies (Barbour et al., 2011; Fogg et al., 2013;
Edwards, Frazer \& Jacoby, 2014; Sandel et al., 2015; Sabido-Itza et
al., 2016) mentioned that some of their organisms were obtained with
pole spears (or other type of harpoon). A single study (de Leon et al.,
2013) did not specify how samples were obtained.

Parameters from models fit to males or females exclusively tend to have
a higher steepness (\emph{i.e.} higher allometric parameter), with mean
\(\pm\) standard deviation values of \(b = 3.27 \pm 0.06\) and
\(b = 3.31 \pm 0.23\) males and females respectively, compared to
parameters from models for pooled genders with a mean \(\pm\) standard
deviation value of \(b = 3.09 \pm 0.22\). In the case of the ponderal
index (\(a\)) and its \(log_{10}\) transformed parameter (\(c\)), values
were higher for parameters for pooled genders. Figure 4 shows the
predicted weights for organisms within the size range of these study
using the 14 parameters previously described.

From all allometric parameters reviewed, those of Edwards, Frazer \&
Jacoby (2014) slightly underestimated the observed biomass, with median
(0.96) and mean (0.98) bellow one. There is no significance in comparing
the parameters estimated in this study since, by definition, they are
the pair of parameters for which residual sum of squares was minimized
with values of median =1.00 and mean =1.03. For all the other studies,
the 95\% confidence interval around the mean fell further away from a
value of 1, indicating overestimation. Predicted to observed biomass
ratios are presented in Figure 5.

\section{Discussion}\label{discussion}

This study provides the first pair of allometric growth parameters
specific to the Central Mexican Caribbean, complementing other studies
performed in Mexican waters in the Alacranes Reef (Aguilar-Perera \&
Quijano-Puerto, 2016) and Xcalak National Park (Sabido-Itza et al.,
2016). By using hand nets instead of spears, we are able to sample a
wider range of sizes often ignored by pole spear samples, allowing us to
include smaller organisms. Estimating parameters by including smaller
organisms ensures better estimatio of weight for smaller sizes. This is
especially important when biomass is calculated from visual census,
where small organisms can be registered. Thus, this study is the first
to provide allometric growth parameters for lionfish in the central
region of the Mexican Caribbean and increases certainty in weight
estimation of small organisms.

The parameters estimated in this study were within the range of studies
in other regions (Barbour et al., 2011; de Leon et al., 2013; Fogg et
al., 2013; Edwards, Frazer \& Jacoby, 2014; Sandel et al., 2015;
Aguilar-Perera \& Quijano-Puerto, 2016; Chin, Aiken \& Buddo, 2016;
Sabido-Itza et al., 2016). As compared to other parameters, the ones
presented here tend to provide lower weight estimates for a same length.
Until about TL = 200 mm, there are no appreciable differences between
the parameters described in this study and those for little Cayman
(Edwards, Frazer \& Jacoby, 2014) and Jamaica (Chin, Aiken \& Buddo,
2016). Yet, for larger organisms (TL \textgreater{} 270 mm) parameters
from Costa Rica (Sandel et al., 2015) and Bonaire (de Leon et al., 2013)
provide similar estimates to those from this study. Conversely, these
same studies tend to overestimate weight -as compared to the ones
reported here- for smaller organisms, likely due to the lack of small
organisms in the samples used to estimate their parameters. When ever
possible, future works should consider the use of hand nets to obtain
the samples not only for studies on weight-at-length, but also diet,
behavior and life history, where length can be an important factor.

There are evident differences in weight-at-length between organisms from
the Caribbean and Gulf of Mexico / North-Western Atlantic. Weight
estimates with parameters from the Gulf of Mexico and North-Western
Atlantic (Barbour et al., 2011; Fogg et al., 2013; Aguilar-Perera \&
Quijano-Puerto, 2016; Sabido-Itza et al., 2016) tend to be higher than
those from the Caribbean (de Leon et al., 2013; Edwards, Frazer \&
Jacoby, 2014; Sandel et al., 2015; Chin, Aiken \& Buddo, 2016), except
for the ones from Xcalak National Park, Mexico (Sabido-Itza et al.,
2016). This indicates that there are differences between lionfish across
the invasion range, pattern that has also been reported for age and
growth parameters (Fogg et al., 2015). These differences can have major
implications in management, especially when estimating biomass available
for harvest or predicting effects on local ecosystems, or evaluating the
effectiveness of removal programs. Using site-specific values provides a
more accurate estimate of fish biomass. Future research should try to
use, to the extent possible, parameters calculated for their region, or
use different parameters to provide upper and lower bounds in their
results. At the same time, this highlights the need for more basic
research. To have a better invasion management, we must perform research
that can describe biologically important information of lionfish
throughout its invasion range (Johnson \& Swenarton, 2016).

While performing the literature review, it was often unclear if
parameters were presented for eq.1 or eq. 3. Sometimes, they were even
mislabeled and yielded senseless results when using the suggested
conversion equation. On some others, parameters were said to be reported
for mm to gr conversions, when they were actually reported as cm to gr
conversions. Perhaps these minor discrepancies can be easily solved by
any researcher, but why should they exist in the first place? It is
important that we report our information in a standard way, making it
readily available for other researchers and managers. In this particular
case, I provide my humble opinion through 5 guidelines to report
allometric parameters:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Be explicit in the methods section. What may seem obvious to you as an
  author -because you have been deeply immersed throughout the process-
  may not be clear to the reader. Specify any transformation performed
  on the data. When using log-transformations, mention the base used to
  transform. Do not assume that ``data were log-transformed'' means
  \(log_{10}(X)\). These assumptions vary across disciplines and
  software and can be a source of confusion. For example, in biology we
  often assume ``log-transformed'' indicates the use of base 10, however
  in R the proper command is \texttt{log10()} and not \texttt{log()},
  which uses base \(e\).
\item
  Use mm and gr to measure TL and TW, respectively. While conversion is
  always possible, we should aim at using standard units to report these
  parameters. If you prefer to use cm to gr conversions, that is
  certainly valid, but make sure to explicitly mention units when
  presenting the parameters.
\item
  Specify the equation for which parameters are presented by including
  an explicit example with the parameters substituted into it, as done
  by some of the papers reviewed (de Leon et al., 2013; Sandel et al.,
  2015; Chin, Aiken \& Buddo, 2016; Sabido-Itza et al., 2016). If
  possible, present the parameters in their exponential (eq. 1) and
  linear (eq. 3) forms.
\item
  Report standard errors and/or confidence intervals around the obtained
  estimates. Given that small changes in \(a\), \(c\), and \(b\) can
  result in important changes in estimated weight, it is important that
  we report uncertainty around each parameter and not just general model
  fit and significance. Reporting uncertainty around parameters allows
  researchers and managers to include upper and lower bounds in their
  predictions.
\item
  Make your data -and code- available. Even if this is not requited by
  the journal or publisher, you can use free cloud data storage services
  or third-party repositories to make your research accessible to
  others. Resources will always be limited and budget will never be
  enough. It is important that we take advantage of open science tools
  that promote the advancement of knowledge and foster collaboration.
  Ultimately, this promotes transparency allows replicability of
  research and advances science.
\end{enumerate}

As final conclusion statements 1) A new pair of allometric growth
parameters are provided for lionfish from the central Mexican Caribbean,
where they exhibit different weight-at-length, as compared to other
regions. 2) Given the differences in weight-at-length patterns through
the invasion range, it is important to use site-specific parameters. 3)
When possible, collect organisms with hand nets to obtain a
representative sample of the population that includes small organisms.
4) Report allometric parameters (and other life history parameters) in a
way that is readily available for non-experts, and make your research
available to advance science.

\section{Aknowledgements}\label{aknowledgements}

I would like to thank thank Nils Van Der Haar and Michael Doodey from
Dive Aventuras as well as Guillermo Lotz-Cador who provided help to
collect samples. I would also like to (anticipatedly) thank the editor
and reviewers, who significantly improved the quality of this work. This
research was partially funded by the ``Consejo Nacional de Ciencias y
Tecnología'' (CONACyT) and the Latin American Fisheries Fellowship
Program.

\clearpage

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-aguilarperera_2016}{}
Aguilar-Perera, A. \& Quijano-Puerto, L. 2016. Relations between fish
length to weight, and otolith length and weight, of the lionfish pterois
volitans in the parque nacional arrecife alacranes, southern gulf of
mexico. \emph{Rev. biol. mar. oceanogr.} 51(2):469--474. DOI:
\href{https://doi.org/10.4067/S0718-19572016000200025}{10.4067/S0718-19572016000200025}.

\hypertarget{ref-aguilarperera_2010}{}
Aguilar-Perera, A. \& Tuz-Sulub, A. 2010. Non-native, invasive red
lionfish (pterois volitans {[}linnaeus, 1758{]}: Scorpaenidae), is first
recorded in the southern gulf of mexico, off the northern yucatan
peninsula, mexico. \emph{AI}. 5(Supplement 1):S9--S12. DOI:
\href{https://doi.org/10.3391/ai.2010.5.S1.003}{10.3391/ai.2010.5.S1.003}.

\hypertarget{ref-albins_2008}{}
Albins, M. \& Hixon, M. 2008. Invasive indo-pacific lionfish pterois
volitans reduce recruitment of atlantic coral-reef fishes. \emph{Mar.
Ecol. Prog. Ser.} 367:233--238. DOI:
\href{https://doi.org/10.3354/meps07620}{10.3354/meps07620}.

\hypertarget{ref-rmarkdown_2017}{}
Allaire, J., Cheng, J., Xie, Y., McPherson, J., Chang, W., Allen, J.,
Wickham, H., Atkins, A., et al. 2017. \emph{Rmarkdown: Dynamic documents
for r}. ed. (nos.). Available:
\url{https://CRAN.R-project.org/package=rmarkdown}.

\hypertarget{ref-andradibrown_2017}{}
Andradi-Brown, D.A., Grey, R., Hendrix, A., Hitchner, D., Hunt, C.L.,
Gress, E., Madej, K., Parry, R.L., et al. 2017. Depth-dependent effects
of culling-do mesophotic lionfish populations undermine current
management? \emph{R Soc Open Sci}. 4(5):170027. DOI:
\href{https://doi.org/10.1098/rsos.170027}{10.1098/rsos.170027}.

\hypertarget{ref-ariasgonzalez_1998}{}
Arias-Gonzalez, J.E. 1998. Trophic models of protected and unprotected
coral reef ecosystems in the south of the mexican caribbean. \emph{J
Fish Biol}. 53(sa):236--255. DOI:
\href{https://doi.org/10.1111/j.1095-8649.1998.tb01030.x}{10.1111/j.1095-8649.1998.tb01030.x}.

\hypertarget{ref-ariasgonzalez_2011}{}
Arias-Gonzalez, J.E., Gonzalez-Gandara, C., Luis Cabrera, J. \&
Christensen, V. 2011. Predicted impact of the invasive lionfish pterois
volitans on the food web of a caribbean coral reef. \emph{Environ Res}.
111(7):917--925. DOI:
\href{https://doi.org/10.1016/j.envres.2011.07.008}{10.1016/j.envres.2011.07.008}.

\hypertarget{ref-barbour_2010}{}
Barbour, A., Montgomery, M., Adamson, A., D?az-Ferguson, E. \& Silliman,
B. 2010. Mangrove use by the invasive lionfish pterois volitans.
\emph{Mar. Ecol. Prog. Ser.} 401:291--294. DOI:
\href{https://doi.org/10.3354/meps08373}{10.3354/meps08373}.

\hypertarget{ref-barbour_2011}{}
Barbour, A.B., Allen, M.S., Frazer, T.K. \& Sherman, K.D. 2011.
Evaluating the potential efficacy of invasive lionfish (pterois
volitans) removals. \emph{PLoS ONE}. 6(5):e19666. DOI:
\href{https://doi.org/10.1371/journal.pone.0019666}{10.1371/journal.pone.0019666}.

\hypertarget{ref-bax_2003}{}
Bax, N., Williamson, A., Aguero, M., Gonzalez, E. \& Geeves, W. 2003.
Marine invasive alien species: A threat to global biodiversity.
\emph{Marine Policy}. 27(4):313--323. DOI:
\href{https://doi.org/10.1016/S0308-597X(03)00041-1}{10.1016/S0308-597X(03)00041-1}.

\hypertarget{ref-betancurr_2011}{}
Betancur-R, R., Hines, A., Acero, A., Orti, G., Wilbur, A. \&
Freshwater, D. 2011. Reconstructing the lionfish invasion: Insights into
greater caribbean biogeography. \emph{J Biogeography}. 38:1281--1293.
DOI:
\href{https://doi.org/10.1111/j.1365-2699.2011.02496.x}{10.1111/j.1365-2699.2011.02496.x}.

\hypertarget{ref-rgdal_2017}{}
Bivand, R., Keitt, T. \& Rowlingson, B. 2017. \emph{Rgdal: Bindings for
the geospatial data abstraction library}. ed. (nos.). Available:
\url{https://CRAN.R-project.org/package=rgdal}.

\hypertarget{ref-bozec_2008}{}
Bozec, Y., Acosta-Gonz\a'alez, G., N\a'uñez-Lara, E. \&
Arias-Gonz\a'alez, J. 2008. Impacts of coastal development on ecosystem
structure and function of yucatan coral reefs, mexico. In
\emph{Proceedings of the 11th international coral reef symposium}. ed.
I.C.R.S. ICRF, Ed. (nos.). Ft. Lauderdale, Florida: 11th International
Coral Reef Symposium.

\hypertarget{ref-chin_2016}{}
Chin, D.A., Aiken, K.A. \& Buddo, D. 2016. Lionfish population density
in discovery bay, jamaica. \emph{International Journal of Scientific \&
Engineering Research}. 7(12):1327--1331.

\hypertarget{ref-cote_2013}{}
Cote, I., Green, S., Morris, J., Akins, J. \& Steinke, D. 2013. Diet
richness of invasive indo-pacific lionfish revealed by dna barcoding.
\emph{Mar. Ecol. Prog. Ser.} 472:249--256. DOI:
\href{https://doi.org/10.3354/meps09992}{10.3354/meps09992}.

\hypertarget{ref-dahl_2014}{}
Dahl, K.A. \& Patterson, W.F. 2014. Habitat-specific density and diet of
rapidly expanding invasive red lionfish, pterois volitans, populations
in the northern gulf of mexico. \emph{PLoS ONE}. 9(8):e105852. DOI:
\href{https://doi.org/10.1371/journal.pone.0105852}{10.1371/journal.pone.0105852}.

\hypertarget{ref-davis_2003}{}
DAVIS, M.A. 2003. Biotic globalization: Does competition from introduced
species threaten biodiversity? \emph{Bioscience}. 53(5):481. DOI:
\href{https://doi.org/10.1641/0006-3568(2003)053\%5B0481:BGDCFI\%5D2.0.CO;2}{10.1641/0006-3568(2003)053{[}0481:BGDCFI{]}2.0.CO;2}.

\hypertarget{ref-deleon_2013}{}
de Leon, R., Vane, K., Bertuol, P., Chamberland, V.C., Simal, F., Imms,
E. \& Vermeij, M.J.A. 2013. Effectiveness of lionfish removal efforts in
the southern caribbean. \emph{Endanger Species Res}. 22(2):175--182.
DOI: \href{https://doi.org/10.3354/esr00542}{10.3354/esr00542}.

\hypertarget{ref-edwards_2014}{}
Edwards, M.A., Frazer, T.K. \& Jacoby, C.A. 2014. Age and growth of
invasive lionfish (pterois spp.) in the caribbean sea, with implications
for management. \emph{BMS}. 90(4):953--966. DOI:
\href{https://doi.org/10.5343/bms.2014.1022}{10.5343/bms.2014.1022}.

\hypertarget{ref-fogg_2015}{}
Fogg, A.Q., Evans, J.T., Ingram JR, G.W., Peterson, M.S. \&
Brown-Peterson, N.J. 2015. Comparing age and growth patterns of invasive
lionfish among three ecoregions of the northern gulf of mexico. In
\emph{Proceedings of the 68 th gulf and caribbean fisheries institute}.
ed. G. GCFI \& C.F. Institute, Eds. (nos.). Panama City: Gulf; Caribbean
Fisheries Institute.

\hypertarget{ref-fogg_2013}{}
Fogg, A.Q., Hoffmayer, E.R., Driggers, W.B., Campbell, M.D., Pellegrin,
G.J. \& Stein, W. 2013. Distribution and length frequency of invasive
lionfish (pterois sp.) in the northern gulf of mexico. \emph{GCR}. 25.
DOI: \href{https://doi.org/10.18785/gcr.2501.08}{10.18785/gcr.2501.08}.

\hypertarget{ref-froese_website_2016}{}
Froese, R. \& Pauly, D. 2016. Available: \url{http://www.fishbase.org/}
{[}2016, December 15{]}.

\hypertarget{ref-green_2012}{}
Green, S.J., Akins, J.L., Maljkovi\a'c, A. \& Côt\a'e, I.M. 2012.
Invasive lionfish drive atlantic coral reef fish declines. \emph{PLoS
ONE}. 7(3):e32596. DOI:
\href{https://doi.org/10.1371/journal.pone.0032596}{10.1371/journal.pone.0032596}.

\hypertarget{ref-grieve_2016}{}
Grieve, B., Curchitser, E. \& Rykaczewski, R. 2016. Range expansion of
the invasive lionfish in the northwest atlantic with climate change.
\emph{Mar. Ecol. Prog. Ser.} 546:225--237. DOI:
\href{https://doi.org/10.3354/meps11638}{10.3354/meps11638}.

\hypertarget{ref-hackerott_2017}{}
Hackerott, S., Valdivia, A., Cox, C.E., Silbiger, N.J. \& Bruno, J.F.
2017. Invasive lionfish had no measurable effect on prey fish community
structure across the belizean barrier reef. \emph{PeerJ}. 5:e3270. DOI:
\href{https://doi.org/10.7717/peerj.3270}{10.7717/peerj.3270}.

\hypertarget{ref-hixon_2016}{}
Hixon, M., Green, S., Albins, M., Akins, J. \& Morris, J. 2016.
Lionfish: A major marine invasion. \emph{Mar. Ecol. Prog. Ser.}
558:161--165. DOI:
\href{https://doi.org/10.3354/meps11909}{10.3354/meps11909}.

\hypertarget{ref-stargazer_2015}{}
Hlavac, M. 2015. \emph{Stargazer: Well-formatted regression and summary
statistics tables}. ed. (nos.). Cambridge, USA: Harvard University.
Available: \url{http://CRAN.R-project.org/package=stargazer}.

\hypertarget{ref-johnson_2016}{}
Johnson, E.G. \& Swenarton, M.K. 2016. Age, growth and population
structure of invasive lionfish (pterois volitans/miles) in northeast
florida using a length-based, age-structured population model.
\emph{PeerJ}. 4:e2730. DOI:
\href{https://doi.org/10.7717/peerj.2730}{10.7717/peerj.2730}.

\hypertarget{ref-johnston_2015}{}
Johnston, M. \& Purkis, S. 2015. A coordinated and sustained
international strategy is required to turn the tide on the atlantic
lionfish invasion. \emph{Mar. Ecol. Prog. Ser.} 533:219--235. DOI:
\href{https://doi.org/10.3354/meps11399}{10.3354/meps11399}.

\hypertarget{ref-jud_2011}{}
Jud, Z., Layman, C., Lee, J. \& Arrington, D. 2011. Recent invasion of a
florida (usa) estuarine system by lionfish pterois volitans / p. miles~.
\emph{Aquat. Biol.} 13(1):21--26. DOI:
\href{https://doi.org/10.3354/ab00351}{10.3354/ab00351}.

\hypertarget{ref-RefManager_2014}{}
McLean, M.W. 2014. \emph{Straightforward bibliography management in r
using the refmanager package}. ed. (nos.). Available:
\url{http://arxiv.org/abs/1403.2036}.

\hypertarget{ref-molnar_2008}{}
Molnar, J.L., Gamboa, R.L., Revenga, C. \& Spalding, M.D. 2008.
Assessing the global threat of invasive species to marine biodiversity.
\emph{Frontiers in Ecology and the Environment}. 6(9):485--492. DOI:
\href{https://doi.org/10.1890/070064}{10.1890/070064}.

\hypertarget{ref-morris_2009}{}
Morris, J.A. \& Akins, J.L. 2009. Feeding ecology of invasive lionfish
(pterois volitans) in the bahamian archipelago. \emph{Environ. Biol.
Fishes}. 86(3):389--398. DOI:
\href{https://doi.org/10.1007/s10641-009-9538-8}{10.1007/s10641-009-9538-8}.

\hypertarget{ref-morris_2011}{}
Morris, J.A., Shertzer, K.W. \& Rice, J.A. 2011. A stage-based matrix
population model of invasive lionfish with implications for control.
\emph{Biol Invasions}. 13(1):7--12. DOI:
\href{https://doi.org/10.1007/s10530-010-9786-8}{10.1007/s10530-010-9786-8}.

\hypertarget{ref-muoz_2011}{}
Muñoz, R., Currin, C. \& Whitfield, P. 2011. Diet of invasive lionfish
on hard bottom reefs of the southeast usa: Insights from stomach
contents and stable isotopes. \emph{Mar. Ecol. Prog. Ser.} 432:181--193.
DOI: \href{https://doi.org/10.3354/meps09154}{10.3354/meps09154}.

\hypertarget{ref-murray_2007}{}
Murray, G. 2007. Constructing paradise: The impacts of big tourism in
the mexican coastal zone. \emph{Coastal Management}. 35(2-3):339--355.
DOI:
\href{https://doi.org/10.1080/08920750601169600}{10.1080/08920750601169600}.

\hypertarget{ref-sp_2017}{}
Pebesma, E.J. \& Bivand, R.S. 2005. Classes and methods for spatial data
in R. \emph{R News}. 5(2):9--13. Available:
\url{https://CRAN.R-project.org/doc/Rnews/}.

\hypertarget{ref-pimentel_2005}{}
Pimentel, D., Zuniga, R. \& Morrison, D. 2005. Update on the
environmental and economic costs associated with alien-invasive species
in the united states. \emph{Ecological Economics}. 52(3):273--288. DOI:
\href{https://doi.org/10.1016/j.ecolecon.2004.10.002}{10.1016/j.ecolecon.2004.10.002}.

\hypertarget{ref-rcore_2017}{}
R Core Team. 2017. \emph{R: A language and environment for statistical
computing}. ed. (nos.). Vienna, Austria: R Foundation for Statistical
Computing. Available: \url{https://www.R-project.org/}.

\hypertarget{ref-broom_2017}{}
Robinson, D. 2017. \emph{Broom: Convert statistical analysis objects
into tidy data frames}. ed. (nos.). Available:
\url{https://CRAN.R-project.org/package=broom}.

\hypertarget{ref-rocha_2015}{}
Rocha, L.A., Rocha, C.R., Baldwin, C.C., Weigt, L.A. \& McField, M.
2015. Invasive lionfish preying on critically endangered reef fish.
\emph{Coral Reefs}. 34(3):803--806. DOI:
\href{https://doi.org/10.1007/s00338-015-1293-z}{10.1007/s00338-015-1293-z}.

\hypertarget{ref-ruizzarate_2004}{}
Ruiz-Zarate, M. \& Arias-Gonzalez, J. 2004. Spatial study of juvenile
corals in the northern region of the mesoamerican barrier reef system
(mbrs). \emph{Coral Reefs}. (September, 9). DOI:
\href{https://doi.org/10.1007/s00338-004-0420-z}{10.1007/s00338-004-0420-z}.

\hypertarget{ref-sabidoitza_2016}{}
Sabido-Itza, M., Medina-Quej, A., De Jesus-Navarrete, A., Gomez-Poot, J.
\& Garcia-Rivas, M. 2016. Uso de la estructura de tallas como evidencia
del establecimiento poblacional del pez le?n pterois volitans
(scorpaeniformes: Scorpaenidae) en el sur del caribe mexicano.
\emph{RBT}. 64(1):353. DOI:
\href{https://doi.org/10.15517/rbt.v64i1.18943}{10.15517/rbt.v64i1.18943}.

\hypertarget{ref-sandel_2015}{}
Sandel, V., Mart\a'inez-Fern\a'andez, D., Wangpraseurt, D. \& Sierra, L.
2015. Ecology and management of the invasive lionfish pterois
volitans/miles complex (perciformes: Scorpaenidae) in southern costa
rica. \emph{Rev Biol Trop}. 63(1):213--221. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/26299126} {[}2017, June 27{]}.

\hypertarget{ref-schofield_2009}{}
Schofield, P. 2009. Geographic extent and chronology of the invasion of
non-native lionfish (pterois volitans {[}linnaeus 1758{]} and p. miles
{[}bennett 1828{]}) in the western north atlantic and caribbean sea.
\emph{AI}. 4(3):473--479. DOI:
\href{https://doi.org/10.3391/ai.2009.4.3.5}{10.3391/ai.2009.4.3.5}.

\hypertarget{ref-schofield_2010}{}
Schofield, P. 2010. Update on geographic spread of invasive lionfishes
(pterois volitans {[}linnaeus, 1758{]} and p. miles {[}bennett, 1828{]})
in the western north atlantic ocean, caribbean sea and gulf of mexico.
\emph{AI}. 5(Supplement 1):S117--S122. DOI:
\href{https://doi.org/10.3391/ai.2010.5.S1.024}{10.3391/ai.2010.5.S1.024}.

\hypertarget{ref-tmap_2017}{}
Tennekes, M. 2017a. \emph{Tmap: Thematic maps}. ed. (nos.). Available:
\url{https://CRAN.R-project.org/package=tmap}.

\hypertarget{ref-tmaptools_2017}{}
Tennekes, M. 2017b. \emph{Tmaptools: Thematic map tools}. ed. (nos.).
Available: \url{https://CRAN.R-project.org/package=tmaptools}.

\hypertarget{ref-usseglio_2017}{}
Usseglio, P., Selwyn, J.D., Downey-Wall, A.M. \& Hogan, J.D. 2017.
Effectiveness of removals of the invasive lionfish: How many dives are
needed to deplete a reef? \emph{PeerJ}. 5:e3043. DOI:
\href{https://doi.org/10.7717/peerj.3043}{10.7717/peerj.3043}.

\hypertarget{ref-valdezmoreno_2012}{}
Valdez-Moreno, M., Quintal-Lizama, C., G\a'omez-Lozano, R. \&
Garc\a'ia-Rivas, M.D.C. 2012. Monitoring an alien invasion: DNA
barcoding and the identification of lionfish and their prey on coral
reefs of the mexican caribbean. \emph{PLoS ONE}. 7(6):e36636. DOI:
\href{https://doi.org/10.1371/journal.pone.0036636}{10.1371/journal.pone.0036636}.

\hypertarget{ref-villaseorderbez_2014}{}
Villaseñor-Derbez, J.C. \& Herrera-P\a'erez, R. 2014. Brief description
of prey selectivity and ontogenetic changes in the diet of the invasive
lionfish pterois volitans (actinopterygii, scorpaenidae) in the mexican
caribbean. \emph{PANAMJAS}. 9(2):131--135.

\hypertarget{ref-tidyverse_2017}{}
Wickham, H. 2017. \emph{Tidyverse: Easily install and load 'tidyverse'
packages}. ed. (nos.). Available:
\url{https://CRAN.R-project.org/package=tidyverse}.

\hypertarget{ref-knitr_2017}{}
Xie, Y. 2017. \emph{Knitr: A general-purpose package for dynamic report
generation in r}. ed. (nos.). Available: \url{http://yihui.name/knitr/}.

\hypertarget{ref-sandwich_2014}{}
Zeileis, A. 2004. Econometric computing with hc and hac covariance
matrix estimators. \emph{Journal of Statistical Software}. 11(10):1--17.
Available: \url{http://www.jstatsoft.org/v11/i10/}.

\hypertarget{ref-lmtest_2002}{}
Zeileis, A. \& Hothorn, T. 2002. Diagnostic checking in regression
relationships. \emph{R News}. 2(3):7--10. Available:
\url{https://CRAN.R-project.org/doc/Rnews/}.

\hypertarget{ref-kableExtra_2017}{}
Zhu, H. 2017. \emph{KableExtra: Construct complex table with 'kable' and
pipe syntax}. ed. (nos.). Available:
\url{https://CRAN.R-project.org/package=kableExtra}.

\clearpage

\clearpage

\section{Tables}\label{tables}

Table 1: Coordinates, minimum, maximum and mean depth (m), and number of
samples for each location.

\begin{tabular}{llllllr}
\toprule
Location & Lat. & Long. & Min. Depth & Max. Depth & Mean Depth & n\\
\midrule
Canones & 20.477 & -87.233 & 15.0 & 31.2 & 21.6 & 11\\
Castillo & 20.496 & -87.220 & 12.5 & 30.5 & 27.5 & 18\\
Cuevitas & 20.478 & -87.244 & 7.4 & 12.8 & 11.2 & 4\\
Islas & 20.490 & -87.228 & 14.0 & 19.4 & 16.7 & 10\\
Paamul & 20.513 & -87.192 & 9.9 & 22.7 & 15.5 & 31\\
\addlinespace
Paraiso & 20.484 & -87.226 & 9.4 & 38.1 & 17.7 & 16\\
Pared & 20.502 & -87.212 & 12.1 & 21.0 & 16.3 & 12\\
Pedregal & 20.507 & -87.204 & 14.4 & 14.9 & 14.7 & 3\\
Santos & 20.493 & -87.222 & 5.7 & 26.6 & 16.2 & 2\\
Tzimin-Ha & 20.393 & -87.307 & 21.2 & 24.6 & 22.9 & 2\\
Total &  &  & 5.7 & 38.1 & 18.6 & 109\\
\bottomrule
\end{tabular}

\clearpage

Table 2: Regression table for the linear model fit between
log10-transformed Total Weight (dependent variable) and Total Length
(independent variable). Numbers in parentheses next to coefficient
estimates indicate heteroskedastic-robust standard errors. The asterisks
(*) indicate the statistical significance. 95\% Confidence intervals are
provided for each parameter.

\begin{table}[!htbp] \centering 
  \caption{} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & log10(TW) \\ 
\hline \\[-1.8ex] 
 c & $-$5.494 (0.083)$^{***}$ \\ 
  b & 3.235 (0.039)$^{***}$ \\ 
 \hline \\[-1.8ex] 
95\% CI for c & (-5.657--5.331) \\ 
85\% CI for b & (3.159-3.311) \\ 
F Statistic & 6928.67*** (df = 1; 107) \\ 
Observations & 109 \\ 
Adjusted R$^{2}$ & 0.976 \\ 
Residual Std. Error & 0.096 (df = 107) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\clearpage

Table 3: Summary of 13 allometric growth parameters available for
lionfish in the invaded range from eight peer-reviewed papers, Fishbase
(Froese \& Pauly, 2016), and this study. All parameters have been
adjusted to convert from millimeters to grams. n = Sample size, a =
scaling parameter for eq. 1, c = y-intercept for eq. 3, b = exponent or
slope for eq. 1 or eq. 3, respectively. The R\textsuperscript{2} column
indicates reported model fit. mDepth and MDepth indicate minimum and
maximum depths (m), respectively, at which organisms were sampled. The
Spear column indicates if the study collected organisms with pole
spears. An asterisk (*) indicates that some portion of organisms were
sampled with pole spears.

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{\begin{tabular}{lllrrrllll}
\toprule
Reference & n & a & c & b & R2 & Sex & mDepth & MDepth & Spear\\
\midrule
Aguilar-Perera \& Quijano-Puerto, 2016 & Both & 472 & 0 & -5.5400 & 3.3000 & 0.95 & 5 & 20 & Yes\\
Aguilar-Perera \& Quijano-Puerto, 2016 & Female & 67 & 0 & -5.9300 & 3.4700 & 0.95 & 5 & 20 & Yes\\
Aguilar-Perera \& Quijano-Puerto, 2016 & Male & 59 & 0 & -5.3800 & 3.2300 & 0.95 & 5 & 20 & Yes\\
Sandel et al., 2015 & Both & 458 & 0 & -4.4400 & 2.8100 & - & - & - & Yes*\\
Chin et al 2016 & Both & 419 & 0 & -4.5600 & 2.8500 & 0.8715 & 18.3 & 18.3 & Yes\\
\addlinespace
Brbour et al 2011 & Both & 774 & 0 & -4.5391 & 2.8900 & - & 27 & 45 & Yes*\\
de Leon et al., 2013 & Both & 1450 & 0 & -4.6411 & 2.8900 & 0.96 & - & - & NA\\
Fogg et al., 2013 & Both & 582 & 0 & -5.8600 & 3.4349 & 0.99 & - & - & Yes*\\
Fogg et al., 2013 & Male & 119 & 0 & -5.5700 & 3.3100 & 0.97 & - & - & Yes*\\
Fogg et al., 2013 & Female & 115 & 0 & -5.1700 & 3.1437 & 0.94 & - & - & Yes*\\
\addlinespace
Edwards et al., 2014 & Both & 1887 & 0 & -5.5229 & 3.2400 & 0.97 & 15 & 30 & Yes*\\
Sabido-Itza et al., 2016 & Both & 2143 & 0 & -5.2828 & 3.1832 & 0.9907 & 0.5 & 57 & Yes*\\
Fishbase & Both & NA & 0 & -5.0293 & 3.0900 & - & - & - & NA\\
This study & Both & 109 & 0 & -5.4941 & 3.2347 & 0.9766 & 5.7 & 38.1 & No\\
\bottomrule
\end{tabular}}
\end{table}

\clearpage

\section{Figures}\label{figures}

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{Map of the study area. The black inset on the left (Yucatan
Peninsula) indicates the location where study sites are distributed. On
the right, circular markers indicate sampling sites and the black
romboid indicates location of Puerto Aventuras, Mexico.}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Kernell density plots for a) Total length (mm) and b) Total
weight (gr) for 109 lionfish sampled in the central Mexican Caribbean.}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{Weight-at-length relationship for 109 lionfish sampled in the
central Mexican Caribbean. Points indicate samples, solid line indicates
curve of best fit.}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{Weight-at-length relationships (n = 14) for eight studies, this
study, and FishBase parameters. Colors indicate studies from which the
parameters were extracted. Solid lines indicate that the fit was
performed for males and females pooled together. Dotted lines indicate
that the regression was performed on females, and dashed lines indicate
it was performed for males.}
\end{figure}

\clearpage

\begin{figure}[htbp]
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{. Box and whiskers plot showing the distribution of predicted
to observed biomass ratios for 14 pairs of allometric parameters. Lines
indicate median values, circles indicate mean values, notches represent
95\% confidence intervals arround the median, lower and upper hinges
correspond to the first and third quartiles, whiskers extend to the
largest and lowes values within 1.5 inter-quartile range of the hinge,
small points represent outliers further away than the whiskers.}
\end{figure}

\end{document}