### 2018-02-24 test.tex

%% Submissions for peer-review must enable line-numbering 
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need 
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file, and
%% the corresponding author info requires v1.2

\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions
% \documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions

\title{Weight-at-length of the invasive lionfish \emph{Pterois volitans}
(Actinopterygii, Scorpaenidae) in the Central Mexican Caribbean, and a
review of allometric growth parameters across the invasion range}

\author[1]{Juan Carlos Villaseñor-Derbez}
\affil[1]{Bren School of Environmental Sciences and Management, University of
California Santa Barbara, Santa Barbara, California, U.S.}
\corrauthor[1]{Juan Carlos Villaseñor-Derbez}{jvillasenor@bren.ucsb.edu}

% \keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
Lionfish (\emph{Pterois volitans/miles}) are an invasive species in the
North-Western Atlantic and the Caribbean. In order to better manage the
invasion, inform lionfish removal programs, and estimate biomass
available for harvest, we must be able to accurately estimate their
total biomass, frequently from length observations. This work compares
length-weight relationships of the invasive lionfish through the
invasion range and reports the length--weight relationship for lionfish
in the Central Mexican Caribbean. A review of 13 length--weight
relationships reported in eight peer-reviewed studies and FishBase is
provided. These parameters were used to identify spatial variation in
weight-at-length. For a given length, parameters from the Caribbean
yielded lower weights than those from the Gulf of Mexico and Atlantic,
indicating that weight-at-length is spatially variable. This highlights
the importance of using site-specific parameters to estimate biomass
from length observations. This study also reports a new pair of
length-weight parameters (\(a = 3.2056 imes 10^{-6}; b = 3.235\)) for
organisms sampled in the Central Mexican Caribbean. Findings from this
work can aid managers and decision makers to better select length-weight
parameters when these are not available for their region of interest.
\end{abstract}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\subsection{R Markdown}\label{r-markdown}

\citep{pimentel_2005}

When you click the \textbf{Knit} button a document will be generated
that includes both content as well as the output of any embedded R code
chunks within the document. You can embed an R code chunk like this:

\begin{verbatim}
##      speed           dist       
##  Min.   : 4.0   Min.   :  2.00  
##  1st Qu.:12.0   1st Qu.: 26.00  
##  Median :15.0   Median : 36.00  
##  Mean   :15.4   Mean   : 42.98  
##  3rd Qu.:19.0   3rd Qu.: 56.00  
##  Max.   :25.0   Max.   :120.00
\end{verbatim}

\subsection{Including Plots}\label{including-plots}

You can also embed plots, for example:

\begin{figure}
\centering
\includegraphics{test_files/figure-latex/pressure-1.pdf}
\caption{\label{fig:pressure}My Caption}
\end{figure}

Note that the \texttt{echo\ =\ FALSE} parameter was added to the code
chunk to prevent printing of the R code that generated the plot.

\bibliography{references}

\end{document}