### 2018-03-20 docs/peerj_template.tex
reorganized the repository

%% Submissions for peer-review must enable line-numbering 
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need 
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file, and
%% the corresponding author info requires v1.2

\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions
% \documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions

\title{$title$}

\author[1]{$author$}
\affil[1]{$institution$}
\corrauthor[1]{$author$}{jvillasenor@bren.ucsb.edu}

% \keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
$abstract$
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

$body$

\end{document}