### 2018-07-13 docs/supplementary/S1Table.Rmd
Changed formatting to submit to BioInv

---
title: Spatial variation in allometric growth of invasive lionfish has management implications
subtitle: "Supplementary Table 1"
author: "Villaseñor-Derbez & Fitzgerald"
output: bookdown::pdf_document2
fontsize: 12pt
---

Coordinates, minimum, maximum and mean depth (m), and number of samples for each location.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.width = 5.5)

options(knitr.table.format = "latex")

windowsFonts(Times=windowsFont("TT Times New Roman"))
```

```{r load packages}
suppressPackageStartupMessages({
  library(stargazer)
  library(sandwich)
  library(lmtest)
  library(broom)
  library(knitr)
  library(tidyverse)
})
```

```{r}
fish <- read_csv(here::here("data", "lionfish_data.csv"), col_types = cols()) %>% 
  group_by(Id, Site, Location, Latitude, Longitude, Depth_m) %>% 
  summarize(Length = 10*mean(TotalLength_cm),
            Weight = mean(TotalWeigth_gr),
            Stomach = mean(StomachLength_cm)) %>% 
  ungroup()
```

```{r table of locations, results = "asis"}
total <- data.frame(Location = "Total", Latitude = "", Longitude = "", minD = min(fish$Depth_m), maxD = max(fish$Depth_m), meanD = formatC(mean(fish$Depth_m), digits = 1, format = "f"), n = 109, stringsAsFactors = F)

fish %>% 
  group_by(Location) %>% 
  summarize(Latitude = formatC(mean(Latitude), digits = 3, format = "f"),
            Longitude = formatC(mean(Longitude), digits = 3, format = "f"),
            minD = formatC(min(Depth_m), digits = 1, format = "f"),
            maxD = formatC(max(Depth_m), digits = 1, format = "f"),
            meanD = formatC(mean(Depth_m), digits = 1, format = "f"),
            n = n()) %>% 
  rbind(total) %>% 
  knitr::kable(row.names = F,
               booktabs = T,
               col.names = c("Location", "Lat.", "Long.", "Min. Depth", "Max. Depth", "Mean Depth", "n"))
```