### 2019-01-30 docs/supplementary/S2Table.Rmd

---
title: Spatial variation in allometric growth of invasive lionfish has management implications
subtitle: "Supplementary Table 2"
author: "Villaseñor-Derbez & Fitzgerald"
output:
  bookdown::pdf_document2:
    toc: no
    keep_tex: yes
fontsize: 12pt
---

Extension of data from Table 1 in the main text. This table includes all the variables collected in each study.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.width = 5.5)

options(knitr.table.format = "latex")

windowsFonts(Times=windowsFont("TT Times New Roman"))
```

```{r load packages}
suppressPackageStartupMessages({
  library(stargazer)
  library(sandwich)
  library(lmtest)
  library(broom)
  library(knitr)
  library(tidyverse)
})
```

```{r load data}
review <- read.csv(here::here("data", "biometric_review.csv"), 
                  stringsAsFactors = F,
                  strip.white = T,
                  blank.lines.skip = T) %>% 
  mutate(c = log10(a),
         Gender2 = case_when(Gender == "Male" ~ "M",
                             Gender == "Female" ~ "F",
                             TRUE ~"B"))
```

```{r}
review %>% 
  filter(!Study == "Fishbase") %>% 
  arrange(desc(Region)) %>% 
  mutate(r2 = ifelse(is.na(r2), "-", r2),
         N = ifelse(is.na(N), "-", N),
         a = formatC((a*1e5), digits = 2),
         tl_range = paste(tl_min, tl_max, sep = " - "),
         tl_range = ifelse(is.na(tl_min), "-", tl_range),
         d_range = paste(minD, maxD, sep = " - "),
         d_range = ifelse(is.na(minD), "-", d_range)) %>% 
  select(Region, Gender2, N, tl_range, a, b, r2, d_range, Years, Spear, Study) %>% 
  knitr::kable(row.names = F,
               digits =  c(0, 0, 0, 0, 2, 2, 2, 0, 0, 0, 0),
               booktabs = T,
               col.names = c("Region",
                             "Sex",
                             "n",
                             "TL range (mm)",
                             "a",
                             "b",
                             "$R^2$",
                             "Depth range (m)",
                             "Years",
                             "Spear",
                             "Study"),
               caption = "Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter, b = exponent. Total Lenth ranges for Barbour et al (2011) and Sandel et al (2015) were estimated from figures since they did not explicitly report these parameters. An asterisk (*) in the Spear column indicates that at least some samples were obtained with pole spears.",
               escape = T) %>% 
  kableExtra::kable_styling(font_size = 10) %>% 
  kableExtra::landscape()
```