### 2019-01-30 docs/supplementary/S2Table.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Spatial variation in allometric growth of invasive lionfish has management implications},
            pdfauthor={Villaseñor-Derbez \& Fitzgerald},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Spatial variation in allometric growth of invasive lionfish has management implications}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Supplementary Table 2}
  \author{Villaseñor-Derbez \& Fitzgerald}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\begin{document}
\maketitle

Extension of data from Table 1 in the main text. This table includes all the variables collected in each study.

\begin{landscape}\begin{table}[t]

\caption{\label{tab:unnamed-chunk-1}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both genders combined (B), a = scaling parameter, b = exponent. Total Lenth ranges for Barbour et al (2011) and Sandel et al (2015) were estimated from figures since they did not explicitly report these parameters. An asterisk (*) in the Spear column indicates that at least some samples were obtained with pole spears.}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}{lllllrlllll}
\toprule
Region & Sex & n & TL range (mm) & a & b & $R^2$ & Depth range (m) & Years & Spear & Study\\
\midrule
Western Atlantic & B & 774 & 75 - 475 & 2.9 & 2.89 & - & 27 - 45 & 2004 - 2009 & Yes* & Barbour et al., 2011\\
Western Atlantic & B & - & - & 0.25 & 3.29 & - & - & NA & NA & Darling et al., 2011\\
GoM & B & 934 & 67 - 377 & 0.21 & 3.34 & 0.98 & - & 2013 - 2014 & Yes & Dahl \& Patterson, 2014\\
GoM & B & 472 & 98 - 358 & 0.29 & 3.30 & 0.95 & 5 - 20 & 2011 - 2013 & Yes & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & F & 67 & 152 - 345 & 0.12 & 3.47 & 0.95 & 5 - 20 & 2011 - 2013 & Yes & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & M & 59 & 151 - 325 & 0.42 & 3.23 & 0.95 & 5 - 20 & 2011 - 2013 & Yes & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & B & 582 & 80 - 385 & 0.14 & 3.43 & 0.99 & - & 2012 & Yes* & Fogg et al., 2013\\
GoM & M & 119 & 126 - 385 & 0.27 & 3.31 & 0.97 & - & 2012 & Yes* & Fogg et al., 2013\\
GoM & F & 115 & 157 - 337 & 0.68 & 3.14 & 0.94 & - & 2012 & Yes* & Fogg et al., 2013\\
Caribbean & B & 458 & 60 - 380 & 3.6 & 2.81 & - & - & 2011 & Yes* & Sandel et al., 2015\\
Caribbean & B & 419 & 93 - 410 & 2.8 & 2.85 & 0.8715 & 18.3 - 18.3 & 2012 - 2014 & Yes & Chin et al., 2016\\
Caribbean & B & 1450 & - & 2.3 & 2.89 & 0.9216 & - & - & - & de Leon et al., 2013\\
Caribbean & B & 1887 & 27 - 391 & 0.3 & 3.24 & 0.97 & 15 - 30 & 2011 & Yes* & Edwards et al., 2014\\
Caribbean & B & 2143 & 25 - 375 & 0.52 & 3.18 & 0.9907 & 0.5 - 57 & 2009-2012 & Yes* & Sabido-Itza et al., 2016\\
Caribbean & B & 227 & 57.09 - 349 & 0.8 & 3.11 & 0.958 & 1 - 30 & 2010 - 2011 & Yes* & Toledo-Hernández et al., 2014\\
Caribbean & B & 449 & 100 - 395 & 0.23 & 3.25 & 0.97 & - & 2012-2013 & Yes & Sabido-Itza et al., 2016b\\
Caribbean & B & 368 & 60 - 380 & 0.32 & 3.19 & 0.98 & - & 2012-2013 & Yes & Sabido-Itza et al., 2016b\\
Caribbean & B & 109 & 34 - 310 & 0.32 & 3.23 & 0.9766 & 5.7 - 38.1 & 2010 & No & This study\\
\bottomrule
\end{tabular}
\end{table}
\end{landscape}

\end{document}