### 2019-02-05 docs/Manuscript.Rmd
Added changes by Sean

---
title: "Spatial variation in allometric growth of invasive lionfish has management implications"
author: "Juan Carlos Villaseñor-Derbez, Sean Fitzgerald"
institution: "Bren School of Environmental Sciences and Management, University of California Santa Barbara, Santa Barbara, California, U.S."
abstract: |
  Lionfish (\textit{Pterois volitans / miles}) are an invasive species in the Western Atlantic and the Caribbean. Improving management of invasive lionfish populations requires accurate total biomass estimates, which depend on accurate estimates of allometric growth, but sedentary species like lionfish often exhibit high levels of spatial variation in life history characteristics. We reviewed 17 published length-weight relationships for lionfish taken throughout their invasive range and found regional differences that led to significant under- and overestimation of weight estimates. The spatial pattern we observed is consistent with findings from other studies focused on genetics or length-at-age. We show that the use of \textit{ex situ} parameters can result in up to a threefold under- or overestimation of total weight, but using parameters from nearby regions substantially reduces this error. These findings can have implications for management in terms of predicting effects on local ecosystems, evaluating the effectiveness of removal programs, or estimating biomass available for harvest.
keywords:
  - Lionfish
  - invasive species
  - length-weight
  - allometric growth
  - regional variations
bibliography: references.bib
output:
  pdf_document:
    keep_tex: yes
    number_sections: no
    template: peerj_template.tex
header-includes:
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \restylefloat{table}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.width = 5.5)

options(knitr.table.format = "latex")

windowsFonts(Times=windowsFont("TT Times New Roman"))
```

```{r load packages}
suppressPackageStartupMessages({
  library(stringr)
  library(agricolae)
  library(magrittr)
  library(stargazer)
  library(sandwich)
  library(lmtest)
  library(broom)
  library(tmap)
  library(rgdal)
  library(sf)
  library(cowplot)
  library(tidyverse)
})
```

\section*{Introduction}

Lionfish (*Pterois volitans/miles* complex) are an invasive species in the Western Atlantic Ocean and Caribbean Sea, likely introduced through release of aquarium-kept organisms \citep{betancurr_2011}. They are the first invasive marine vertebrates established along these coasts \citep{schofield_2009,schofield_2010,sabidoitza_2016} and their presence has been labeled as a major marine invasion because they threaten local biodiversity, spread rapidly, and are difficult to manage \citep{hixon_2016}. Lionfish have established invasive populations in coral reefs, estuaries, mangroves, hard-bottomed areas, and mesophotic reefs \citep{barbour_2010,jud_2011,muoz_2011,claydon_2012,andradibrown_2017,gress_2017}.

A substantial amount of research describes lionfish impacts throughout its invaded range. A meta-analysis by \citet{peake_2018} showed that invasive lionfish prey on at least 167 different species across the tropical and temperate Western Atlantic. Their feeding behavior and high consumption rates can reduce recruitment and population sizes of native reef-fish species, and can further endanger reef fish \citep[][but see \citealt{hackerott_2017} for a counterexample]{green_2012,rocha_2015}. For example, field experiments showed that lionfish establishment in the Bahamas led to reduced recruitment of native fishes by nearly 80% over a five-week period \citep{albins_2008}, and prey fish biomass declined by 65% over two years as lionfish biomass increased along Bahamian coral reefs \citep{green_2012}. However, trophic impacts of lionfish can be minimized if their local biomass is controlled by culling \citep{ariasgonzalez_2011}.

Governments and non-profit organizations have sought to reduce lionfish densities through removal programs and incentivizing its consumption \citep{chin_2016}. In some cases, these have shown to significantly reduce -- but not quite eliminate -- lionfish abundances at local scales \citep{deleon_2013,sandel_2015}. Complete eradication of lionfish through fishing is unlikely because of their rapid recovery rates and ongoing recruitment to shallow-water areas from persistent populations in mesophotic ecosystems \citep{barbour_2011,andradibrown_2017}. However, promoting lionfish consumption might create a level of demand capable of incentivizing a stable fishery while controlling shallow-water populations, thus creating alternative livelihoods and avoiding further negative effects to local biota.

The feasibility of establishing fisheries through lionfish removal programs has been extensively evaluated through field observations and empirical modeling \citep{barbour_2011,morris_2011,deleon_2013,johnston_2015,sandel_2015,usseglio_2017}. Determining the feasibility of such initiatives requires modeling the change in biomass in response to changes in fishing mortality (*i.e.* culling). A common way to model this is via length-structured population models, where fish lengths are converted to weight to calculate total biomass \citep{barbour_2011,cote_2014,andradibrown_2017}. The allometric length-weight relationship is thus an essential component of these models, but this relationship can vary across regions as a response to biotic and abiotic conditions \citep{johnson_2016}.

Outcomes of previous studies suggest lionfish are likely to exhibit spatial heterogeneity in the length-weight relationship for two main reasons. First, fish with sedentary behavior such as lionfish often exhibit high levels of spatial variation in important life history characteristics such as growth or natural mortality rates \citep{gunderson_2008,hutchinson_2008,wilson_2012,guan_2013}. In fact, culling programs are effective in reducing local adult populations largely because lionfish have high levels of site fidelity and small home ranges \citep{Fishelson_1997,kochzius_2005,jud_2012,cote_2014}. Second, genetic analysis of lionfish suggests biological differences due to the existence of two genetically distinct invasive subpopulations between the Western Atlantic and the Caribbean \citep{betancurr_2011}. Site-specific studies that calculate the length-weight relationship of lionfish report variable estimates, and these differences may be increasingly important when estimating the potential effectiveness of lionfish culling programs \citep{barbour_2011,morris_2011,cote_2014,johnston_2015}. However, the influence of using *ex situ* parameters when estimating the length-weight relationship remains unexplored.

Our objective was to quantify the magnitude of error caused by using *ex situ* parameter values when estimating lionfish weight from length observations. In this study, we calculated and reported the first length-weight relationship for lionfish in the central Mexican Caribbean (n = 109). We then estimated lionfish weight in this area using previously published length-weight relationships for lionfish populations in North Carolina, the Northern and Southern Gulf of Mexico, the Southern Mexican Caribbean, the Bahamas, Little Cayman, Jamaica, Bonaire, Puerto Rico, and Costa Rica. By comparing these weight estimates to the *in situ* length-weight observations, we showed that using *ex situ* parameter values resulted in up to a threefold under- or overestimation of lionfish weight.

<!-- \citep{barbour_2011,darling_2011,deleon_2013,fogg_2013,dahl_2014,edwards_2014,toledohernndez_2014,sandel_2015,aguilarperera_2016,sabidoitza_2016,sabidoitz_2016,chin_2016} -->

<!-- I don't like how the citations cut this paragraph, which is one of the most important ones. I don't think we have to explicitly cite them here, since they are cited in the table and basically all figures, as well as in instances when we touch on each region. If you agree, let's delete them. Otherwise, we can put them right after Costa Rica / before "... By comparing these...". -->

\section*{Methods}

```{r load fish data, include = F}
fish <- read_csv(here::here("data", "lionfish_data.csv"), col_types = cols()) %>%
  group_by(Id, Site, Location, Latitude, Longitude, Depth_m) %>% 
  summarize(Length = 10*mean(TotalLength_cm),
            Weight = mean(TotalWeigth_gr),
            Stomach = mean(StomachLength_cm)) %>% 
  ungroup()

review <- read.csv(here::here("data", "biometric_review.csv"), 
                  stringsAsFactors = F,
                  strip.white = T,
                  blank.lines.skip = T) %>% 
  mutate(c = log10(a),
         Gender2 = case_when(Gender == "Male" ~ "M",
                             Gender == "Female" ~ "F",
                             TRUE ~"B"),
         tl_min = ifelse(is.na(tl_min), 34, tl_min),
         tl_max = ifelse(is.na(tl_max), 310, tl_max))
```

We reviewed 12 published studies and obtained 17 length-weight relationships for the Western Atlantic (n = 2), Gulf of Mexico (n = 7), and Caribbean (n = 8, Table \ref{tab:all_params}, Fig \ref{fig:map}). We collected information on sex differentiation, location, length and depth ranges, and sampling methods from each study when available. Only two studies reported parameters for each sex \citep{aguilarperera_2016,fogg_2013}, so we assumed data from both sexes were combined in a study if sex was unspecified. Reviewed studies presented information for organisms ranging from `r min(review$tl_min, na.rm = T)`-`r max(review$tl_max, na.rm = T)` mm in Total Length ($TL$) and were obtained at depths between `r min(review$minD, na.rm = T)` m and `r max(review$maxD, na.rm = T)` m. Four studies explicitly stated that their organisms were sampled with pole spears \citep{dahl_2014,aguilarperera_2016,chin_2016,sabidoitz_2016}, and six studies mentioned that some of their organisms were obtained with pole spears (or other type of harpoon) but also hand-held nets or fish traps \citep{barbour_2011,fogg_2013,edwards_2014,toledohernndez_2014,sandel_2015,sabidoitza_2016}. Two studies did not specify how organisms were sampled \citep{darling_2011,deleon_2013}.

We also used data from 109 lionfish sampled by \citet{villaseorderbez_2014}, who used hand nets and numbered bottles to collect organisms from 10 sampling sites along the central Mexican Caribbean coast in the Summer of 2010 (Supplementary Table 1). Sampling locations included wall and carpet reefs at depths between 5.7 m and 38.1 m. The use of hand nets prevented any weight loss due to bleeding and allowed better representation of small sizes by avoiding gear selectivity. Organisms were euthanized via pithing and Total Length (TL; mm) and Total Weight (TW; g) were recorded.

\clearpage

```{r}

proj <- c("+proj=longlat +datum=NAD27")

# Load coastline for Mexico and World

data(World)
W <- st_transform(World, proj)

# Map with SF
fig_map <- ggplot() +
  geom_sf(data = W) +
  geom_point(data = review %>%
      mutate(N = ifelse(is.na(N), 1, N)),
    aes(x = lon, y = lat, fill = b, size = N), shape = 21) +
  theme_classic() +
  scale_fill_gradientn(colors = colorRamps::matlab.like(20)) +
  theme(legend.background = element_blank(),
        panel.grid.major = element_line(colour = "transparent")) +
  lims(x = c(-100, -60), y = c(9, 36)) +
  labs(x = "Longitude", y = "Latitude") +
  geom_text(aes(x = -70, y = 30), label = "Western Atlantic") +
  geom_text(aes(x = -90, y = 25), label = "Gulf of Mexico") +
  geom_text(aes(x = -75, y = 15), label = "Caribbean")

```

```{r map, fig.cap = "\\label{fig:map}Locations where allometric growth parameters of lionfish (*Pterois spp*) have been reported. Circle sizes indicate sample size from each study, colors indicate the $b$ coefficient from Eq. \\ref{eq:allometric}.", fig.height = 3.5}
fig_map
```

The weight-at-length relationship for lionfish in the central Mexican Caribbean was calculated with the allometric growth function

\begin{equation}
\label{eq:allometric}
TW = aTL^b
\end{equation}

where $a$ is the ponderal index and $b$ is the scaling exponent or allometric parameter. We linearized the equation using $log_{10}$-transformation and estimated the coefficients using an Ordinary Least Squares Regression with a heteroskedastic-robust standard error correction \citep{zeileis_2004}. Coefficients were tested with a two-tailed Student's t-test, and the significance of the regression was corroborated with an F-test.

Some of the reviewed studies inconsistently defined $a$ as either the ponderal index from Eq. \ref{eq:allometric} or the y-intercept from the linearized log-transformed equation. Other studies incorrectly reported parameters as mm-to-g conversions when they were in fact cm-to-g conversions. We standardized each study by converting coefficients and report all parameters as TL (mm) to TW (g) conversions. Locations where allometric studies have been performed are shown in Figure \ref{fig:map} and summarized in Table \ref{tab:all_params}.

We obtained a total of 18 parameter pairs by combining length-weight parameters extracted from the literature and the additional pair calculated here. Recall that the objective of this study is not to describe variations between populations, but rather to estimate how the use of *ex situ* parameters influences the accuracy of weight estimates for lionfish, using the central Mexican Caribbean as a case study. Using each of the 18 parameter pairs, we estimated $TW$ from the $TL$ observations collected in the central Mexican Caribbean (n = 109, with $TL$ ranging from 34 mm to 310 mm) and divided predicted weights by known observed weights to obtain a simple measure of over- or underestimation. Difference in mean weight ratios were tested with an analysis of covariance (ANCOVA), where the sources of variation were the study and $TL$. Ratios were logit-transformed prior to analysis, and a *post-hoc* Tukey's test was used to identify groups where mean ratios did not differ. All analyses were performed in R version `r paste(version$major, version$minor, sep = ".")` \citep{rcore_2018}. Raw data and code used in this work are available on github at github.com/jcvdav/lionfish_biometry.

```{r}
review %>% 
  select(Region, Gender2, N, a, b, r2, Study) %>% 
  filter(!Study == "Fishbase") %>% 
  arrange(desc(Region)) %>% 
  mutate(r2 = formatC(r2, digits = 2),
         r2 = ifelse(r2 == " NA", "-", r2),
         N = ifelse(is.na(N), "-", N),
         a = formatC((a*1e5), digits = 2)) %>% 
  knitr::kable(row.names = F,
               digits =  c(0, 0, 0, 0, 2, 2),
               booktabs = T,
               col.names = c("Region", "Sex", "n", "a", "b", "$R^2$", "Reference"),
               caption = "\\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both sexes combined (B), a = scaling parameter (presented in $\\times 10^{-5}$), b = exponent.")
```

```{r model fitting}
model <- lm(log10(Weight) ~ log10(Length), data = fish)

TidyModel <- model%>%
  lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC1")) %>%
  broom::tidy()

robust_se <- TidyModel$std.error %>%
  set_names(TidyModel$term)
```

```{r testing coefficients}
test_intercept <- 2*pt(abs(coef(model)[[1]]/robust_se[[1]]), df = 107, lower.tail = F)

test_slope <- 2*pt(abs(coef(model)[[2]]/robust_se[[2]]), df = 107, lower.tail = F)

test_slope_3 <- 2*pt(abs((coef(model)[[2]]-3)/robust_se[[2]]), df = 107, lower.tail = F)

c <- coefficients(model)[[1]]
a <- formatC(10^c, digits = 4)
b <- coefficients(model)[[2]]
bs <- formatC(b, digits = 4)
```

```{r}
Predicted_biomass <- review %>% 
  mutate(Length = list(fish$Length),
         Weight = list(fish$Weight)) %>% 
  unnest() %>% 
  mutate(pWeight = a*(Length^b),
    Biomass_ratio = pWeight / Weight,
    logit_biomass_ratio = car::logit(Biomass_ratio))

anova <- aov(Biomass_ratio ~ Study_id, data = Predicted_biomass)
ancova <- aov(logit_biomass_ratio ~ Study_id + Length, data = Predicted_biomass)

groups <- HSD.test(y = ancova, trt = "Study_id")

Rsq <- formatC(summary(model)$r.sq, digits = 3, format = "f")
Fstring <- waldtest(model, vcov = vcovHC(model, type = "HC1")) %>% 
  tidy()

Fstring <- paste0("F(df = 1; ", nobs(model)-2, ") = ", formatC(Fstring$statistic[2], digits = 2, format = "f"))
```

\section*{Results}

The length-weight relationship for organisms from the central Mexican Caribbean resulted in coefficient values of $a = `r a`$ and $b = `r bs`$ ($R^2 = `r Rsq`$, $F_{1, 107} = 6928.67; p < 0.001$). The allometric factor ($b$) was significantly different from $b = 3$ ($t_{107} = `r formatC((coef(model)[[2]]-3)/robust_se[[2]], digits = 2, format = "f")`; p<0.001$) corroborating that lionfish present allometric growth. The length-weight coefficients estimated here were within the range identified by studies from other regions (Table \ref{tab:all_params}). Figure \ref{fig:l-w-carib} shows the relationship between TL and TW for this region.

```{r}
TidyModel <- model %>%
  lmtest::coeftest(vcov = sandwich::vcovHC(., type = "HC1")) %>%
  broom::tidy()

robust_se <- TidyModel$std.error %>%
  set_names(TidyModel$term)
wald_results <- waldtest(model, vcov = vcovHC(model, type = "HC1"))
f_string <- c("F Statistic", paste0(formatC(wald_results$F[2], digits = 2, format = "f"), "*** (df = 1; ", dim(fish)[1]-2,")"))
```

```{r}
get_eqn <- function(model){
  substitute(italic(TW) == a %*% "10"^"-6" ~ italic(TL)^b,
             list(a = "3.2056",
                  b = "3.2347")) %>% 
    as.expression() %>% 
    as.character()
}

get_eqn2 <- function(model){
  substitute(italic(log[10](TW)) == b %*% italic(log[10](TL)) ~ a,
             list(a = "-5.4940",
                  b = "3.2347")) %>% 
    as.expression() %>% 
    as.character()
}

get_rsq <- function(model){
  substitute(italic(R)^2~"="~r2,
             list(r2 = formatC(summary(model)$r.squared, digits = 3))) %>% 
    as.expression() %>% 
    as.character()
}
```

```{r}
fit <- ggplot(data = fish, mapping = aes(x = Length, y = Weight)) +
  geom_point(size = 3, shape = 21, fill = "steelblue", color = "black") +
  theme_cowplot() +
  labs(x = "Total length (mm)", y = "Total Weight (g)") +
  theme(text = element_text(family = "Times")) +
  scale_color_brewer(palette = "Set3") +
  theme(text=element_text(family = "Times")) +
  geom_smooth(method = "lm", formula = y ~ I(c*x^b), se = F, color = "black", linetype = "dashed") +
  annotate(geom = "text",
            x = 125, y = 390,
            label = get_eqn(model),
            parse = T,
            size = 4,
            family="Times",
            face="italic") +
  annotate(geom = "text",
            x = 125, y = 350,
            label = get_eqn2(model),
            parse = T,
            size = 4,
            family="Times",
            face="italic") +
  annotate(geom = "text",
            x = 125, y = 310, 
            label = get_rsq(model),
            parse = T,
            size = 4,
            family="Times",
            face="italic")

fig_l_w_carib <- ggExtra::ggMarginal(p = fit,
                                     type = "density",
                                     margins = "both",
                                     fill = "steelblue",
                                     color = "black",
                                     alpha = 0.5)
```

```{r fit1,  fig.height = 3.5, fig.pos = "H", fig.cap = "\\label{fig:l-w-carib}Length-weight relationship for 109 lionfish sampled in the central Mexican Caribbean. Points indicate samples, dashed black line indicates curve of best fit, marginal plots represent the density distribution of each variable."}
fig_l_w_carib
```

```{r, results = "asis", eval = F}
stargazer(model,
          dep.var.caption = "$log_{10}(TW)$",
          dep.var.labels.include = F,
          header = F,
          se = list(robust_se),
          type = "latex",
          t.auto = T,
          p.auto = T,
          intercept.bottom = F,
          single.row = T,
          report = "vcs*",
          omit.stat = c("f", "rsq"),
          add.lines = list(f_string),
          covariate.labels = c("c", "b"),
          dep.var.labels = "log10(TW)", title = "\\label{tab:reg_table}Coefficients of the linear model fit to Eq 1. Numbers in parenthesees represent heteroskedastic-robust standard errors.")
```

There were significant differences in our predicted weights for the central Mexican Caribbean when using each of the different pairs of parameters ($F_{17, 1943} = 24.96; p < 0.001$). The lowest weight estimates for the observed lengths resulted from using the allometric parameters from Banco Chinchorro in the Caribbean, with mean $\pm$ SD of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{sabidoitz_2016}. In contrast, the highest weight estimates came from the Western Atlantic with `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(pWeight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(pWeight) %>% formatC(digits = 2, format = "f")` g \citep{barbour_2011}. To put this in context, true observed weights have a mean of `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Weight) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Weight) %>% formatC(digits = 2, format = "f")` g. Weights predicted from these extreme parameters correspond to predicted-to-observed weight ratios of `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Sabido2BC") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` and `r filter(Predicted_biomass, Study_id == "Barbour") %$% mean(Biomass_ratio) %>% formatC(digits = 2, format = "f")` $\pm$ `r filter(Predicted_biomass, Study_id == "Barbour") %$% sd(Biomass_ratio) %>% formatC(digits = 2, format = "f")` (mean $\pm$ SD), respectively. The largest under- and overestimations resulted in ratios of `r min(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")` and `r max(Predicted_biomass$Biomass_ratio) %>% formatC(digits = 2, format = "f")` of the actual observed weight, indicating that *ex situ* parameters can result in highly erroneous weight estimates.

Tukey's *post-hoc* test suggests that weight ratios for the central Mexican Caribbean were not different from those obtained with parameters from Little Cayman, the Bahamas, and some sites in the Gulf of Mexico (Tukey's HSD $p > 0.05$). All weight estimates using parameters from the Gulf of Mexico and Western Atlantic were higher than observed values, and only parameters from the Caribbean produced weights smaller than observed (Fig \ref{fig:bio_ratio}). The regional average ($\pm$ SD) of predicted-to-observed weight ratios from these three regions were 1.24 $\pm$ 0.309, 1.41 $\pm$ 0.523, and 1.20 $\pm$ 0.423 for the Gulf of Mexico, Western Atlantic, and Caribbean, respectively. This suggests that the smallest errors are observed when using parameters from other locations in the Caribbean. Predicted-to-observed weight ratios are presented in Figure \ref{fig:bio_ratio}.

```{r}
This <- data.frame(Length = c(34:310)) %>% 
  mutate(Weight = 10^(coefficients(model)[1] + coefficients(model)[2]*log10(Length)))

fig_all_allo <- review %>% 
  filter(!Study == "This study") %>%
  mutate(Length = list(seq(1, 400))) %>% 
  unnest() %>% 
  filter(Length > tl_min & Length < tl_max) %>% 
  mutate(Weight = a*(Length^b)) %>%
  group_by(Length) %>% 
  mutate(min = min(Weight),
         max = max(Weight),
         Region = case_when(Region == "Western Atlantic" ~ "WA",
                            Region == "Caribbean" ~ "Ca",
                            T ~ "GoM"),
         Study_Region = paste0(Study, " (", Region, ")")) %>% 
  ungroup() %>% 
  ggplot(aes(x = Length)) +
  geom_line(aes(y = Weight, color = Study_Region,
                linetype = Gender, group = Study_id), size = 1) +
  geom_line(data = This, aes(x = Length, y = Weight),
            size = 1.2, color = "black", linetype = "dashed") +
  labs(x = "Total length (mm)", y = "Total Weight (g)") +
  theme_cowplot() +
  theme(text = element_text(family = "Times"),
        legend.justification=c(0,1),
        legend.position=c(0.01, 1.05),
        legend.background = element_rect(fill = "transparent"),
        legend.box = "horizontal",
        legend.text = element_text(size = 9),
        legend.title = element_text(size = 10)) +
  guides(color = guide_legend(order = 1),
         linetype = guide_legend(order = 0, ncol = 1, override.aes = list(size = 0.5)),
         size = "none") +
  scale_color_brewer(palette = "Paired", direction = -1) +
  scale_linetype_manual(values=c("solid", "dashed", "dotted"), name = "Sex") +
  scale_size_manual(values = c(1, 1.5))
```

```{r fit2, fig.height = 3.8, fig.pos = "H", fig.cap = "\\label{fig:all_allo}Length-weight relationships (n = 18) for 12 studies and this study. The curves are shown for the range of lengths reported in each study (See Supplementary Table 2); when ranges were not present, we use the ones found in this study (34 mm - 310 mm). Colors indicate studies from which the parameters were extracted. Dotted, dashed, and solid lines show models for males, females, and combined sexes, respectively. Letters in parentheses indicate if the study comes from the Gulf of Mexico (GoM), Western Atlantic (WA), or Caribbean (Ca). The dashed black line represents the relationship estimated in this study. There are two solid green lines for Sabido-Itza et al, 2016b, one for each of the two sites for which they report parameters. A log-log version of this figure is presented in Figure S4."}
fig_all_allo
```

```{r}
groups$groups%<>%
  mutate(Study_id = rownames(.)) %>% 
  select(-logit_biomass_ratio)

Predicted_biomass %<>% 
  left_join(groups$groups, by = "Study_id") %>% 
  mutate(groups = as.character(groups),
         groups = ifelse(Length == 34, groups, ""))

fig_bio_ratio <- Predicted_biomass %>% 
  mutate(Study = paste0(Study, " (", Gender_site, ")"),
         Region = case_when(Region == "Western Atlantic" ~ "WA",
                            Region == "Caribbean" ~ "Ca",
                            T ~ "GoM")) %>%
  transform(Study = reorder(Study, -Biomass_ratio),
            Study_id = reorder(Study_id, -Biomass_ratio)) %>%
  rename(Sex = Gender) %>% 
  ggplot(aes(x = Study, y = Biomass_ratio, group = Study_id)) +
  geom_violin(aes(fill = Region), alpha = 0.8) +
  stat_summary(geom = "point",
               fun.y = "median",
               size = 2,
               color = "black",
               fill = "steelblue",
               shape = 21) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(y = "Weight ratio (Predicted/Observed)", x = "") +
  coord_flip() +
  theme_cowplot() +
  theme(text = element_text(family = "Times")) +
  scale_fill_brewer(palette = "Set1") +
  theme(text = element_text(family = "Times"),
        legend.position = "top",
        legend.text = element_text(size = 10),
        legend.background = element_rect(fill = "transparent"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12)) +
  geom_text(aes(y = 3.5, label = groups), family = "Times", fontface = "plain") +
  scale_y_continuous(limits = c(NA, 3.7)) +
  guides(fill = guide_legend(title.position = "top"))
```

```{r pred_obs, fig.height = 5.2, fig.width = 5.5, fig.cap = "\\label{fig:bio_ratio}Violin plot of predicted-to-observed weight ratios for 18 pairs of allometric parameters. Blue circles indicate median values and like letters indicate values that do not differ significantly. For Sabido-Itza et al, 2016b, BC and PNAX make reference to Banco Chinchorro and Parque Nacional Arrecifes de Xcalak, two sites for which they report parameters."}
fig_bio_ratio
```

\clearpage

\section*{Discussion}

Our results suggest that lionfish exhibit highly variable, spatially heterogeneous allometric relationships across their invaded range, and that this variation is relevant for managing invasions. Moreover, we show that the use of *ex situ* parameter values may lead to highly biased weight estimates. Our comparison of observed weights to those predicted with locally-informed parameters and *ex situ* parameters showed that weight can be overestimated by more than threefold, highlighting the need to use local information. Here we discuss the implications of our findings, possible shortcommings in our analyses, and highlight potential future research directions.

Differences in length-weight relationships have traditionally been highlighted as potential pitfalls to fishery management. For example, \citet{wilson_2012} show that small-scale variations in length-at-age and fishing mortality in other Scorpaeniformes translate to differential landings, effort, and catch per unit effort in the live fish fishery of California, and that these differences must be taken into account in management plans. The lionfish case poses the opposite scenario, where the manager desires to eradicate the species. To accurately gauge both the effectiveness of lionfish removal efforts and the resources needed to successfully manage an invasion, we must acknowledge and understand regional biological differences in important variables such as allometric growth parameters.

We detected substantial differences in weight-at-length between organisms from the Caribbean, Gulf of Mexico, and Western Atlantic. Groupings of predicted-to-observed weight ratios identified in our *post hoc* testing aligned with the spatial distribution of the examined studies, suggesting that these differences may be mediated by space. These regional allometric differences mirror similar patterns in length-at-age of lionfish across both their invaded and native regions \citep{pusack_2016}. Variation may be driven by genetics or by organisms' exposure to distinct environmental conditions. For example, \citet{betancurr_2011} used mitochondrial DNA to demonstrate the existence of two distinct population groups, identified as the “Caribbean group” and “Northern Group”, and \citet{fogg_2015} alternatively suggested that length-at-age differences may be driven by the environment.

We might be inclined to attribute all variation in the lionfish length-weight relationship to the spatial origin of these parameters. However, samples from the 12 studies included here were not only collected in different locations, but also at different points in time and across different depth and size ranges (See Supplementary Table 2 for an extended version of Table 1 that incorporates these variables when available). While we could not evaluate how these factors influence previous estimates without the necessary raw data from every study, we still show that the lack of locally-calculated parameters can induce significant bias when calculating weight from length observations. The magnitude of the bias discovered in this study and our lack of understanding the sources driving spatial variation for lionfish highlights the need to simultaneously collect length-weight information across the invaded range to test for spatially-induced patterns and to link these findings to previously suggested environmental and genetic structures. Such an endeavor would provide insight into lionfish biology and better inform management.

Applying parameters estimates to lengths outside the range of lengths originally used to estimate the parameters may also induce error. Our smallest observed organism was 34 mm in $TL$, and only two studies estimated parametrs with smaller organisms \citep{edwards_2014,sabidoitza_2016}. However, our largest organism had a $TL$ of 310 mm, which is well within the range of all other studies (the next smallest maximum length was 325 mm; See Supplementary Table 2). Due to the power function describing the allometric relationship (*i.e.* Eq. 1), the error is larger when extrapolation is done for lengths that are larger than the maximum length used to estimate the parameters.

<!-- I modified the paragraph above a bit, and think its a bit more clear, but please change it if you think there are better ways to explain it. I think the central idea is that we do not extrapolate from other predictions, so we are safe in that sense. -->

The results presented here have key implications for management. For example, \citet{edwards_2014} simulated a lionfish culling program under two scenarios, one using length-at-age and length-to-weight parameters from North Carolina and one using parameters from Little Cayman. Their results show that using different parameters caused up to a four-year difference in the time required for the simulated lionfish population to recover to 90% of its initial biomass after removals ceased. Here, we show that using one set of length-weight parameters versus another for a given length can result in more than a threefold under- or overestimation of total weight. These differences become especially important when allocating resources for lionfish removal programs, incentivizing lionfish fisheries as a source of alternative livelihoods, or estimating ecosystem impacts. Research efforts focused on invasive lionfish populations need to use parameters calculated for their region to the extent possible, or use different sets of parameters that provide appropriate upper and lower bounds in their results.

\section*{Acknowledgements}

We thank Nils Van Der Haar and Michael Doodey from Dive Aventuras as well as Guillermo Lotz-Cador who provided help to collect samples. We are grateful for comments raised by the editor and two anonymous reviewers, which significantly increased the quality of this work.

Conflict of Interest: The authors declare that they have no conflict of interest.

\bibliography{references}