### 2019-02-05 docs/Manuscript.tex

%% Submissions for peer-review must enable line-numbering 
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need 
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file, and
%% the corresponding author info requires v1.2

\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions
% \documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions

\title{Spatial variation in allometric growth of invasive lionfish has
management implications}

\author[1]{Juan Carlos Villaseñor-Derbez}
\author[1]{Sean Fitzgerald}
\affil[1]{Bren School of Environmental Sciences and Management, University of California
  Santa Barbara, Santa Barbara, California, USA}
\corrauthor[1]{Juan Carlos Villaseñor-Derbez}{juancarlos@ucsb.edu}

\keywords{Lionfish, invasive species, length-weight, allometric, regional variations}

\begin{abstract}
Lionfish (\textit{Pterois volitans / miles}) are an invasive species in
the Western Atlantic and the Caribbean. Improving management of invasive
lionfish populations requires accurate total biomass estimates, which
depend on accurate estimates of allometric growth, but sedentary species
like lionfish often exhibit high levels of spatial variation in life
history characteristics. We reviewed 17 published length-weight
relationships for lionfish taken throughout their invasive range and
found regional differences that led to significant under- and
overestimation of weight estimates. The spatial pattern we observed is
consistent with findings from other studies focused on genetics or
length-at-age. We show that the use of \textit{ex situ} parameters can
result in up to a threefold under- or overestimation of total weight,
but using parameters from nearby regions substantially reduces this
error. These findings can have implications for management in terms of
predicting effects on local ecosystems, evaluating the effectiveness of
removal programs, or estimating biomass available for harvest.
\end{abstract}

\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}

Lionfish (\emph{Pterois volitans/miles} complex) are an invasive species
in the Western Atlantic Ocean and Caribbean Sea, likely introduced
through release of aquarium-kept organisms \citep{betancurr_2011}. They
are the first invasive marine vertebrates established along these coasts
\citep{schofield_2009,schofield_2010,sabidoitza_2016} and their presence
has been labeled as a major marine invasion because they threaten local
biodiversity, spread rapidly, and are difficult to manage
\citep{hixon_2016}. Lionfish have established invasive populations in
coral reefs, estuaries, mangroves, hard-bottomed areas, and mesophotic
reefs
\citep{barbour_2010,jud_2011,muoz_2011,claydon_2012,andradibrown_2017,gress_2017}.

A substantial amount of research describes lionfish impacts throughout
its invaded range. A meta-analysis by \citet{peake_2018} showed that
invasive lionfish prey on at least 167 different species across the
tropical and temperate Western Atlantic. Their feeding behavior and high
consumption rates can reduce recruitment and population sizes of native
reef-fish species, and can further endanger reef fish
\citep[][but see \citealt{hackerott_2017} for a counterexample]{green_2012,rocha_2015}.
For example, field experiments showed that lionfish establishment in the
Bahamas led to reduced recruitment of native fishes by nearly 80\% over
a five-week period \citep{albins_2008}, and prey fish biomass declined
by 65\% over two years as lionfish biomass increased along Bahamian
coral reefs \citep{green_2012}. However, trophic impacts of lionfish can
be minimized if their local biomass is controlled by culling
\citep{ariasgonzalez_2011}.

Governments and non-profit organizations have sought to reduce lionfish
densities through removal programs and incentivizing its consumption
\citep{chin_2016}. In some cases, these have shown to significantly
reduce -- but not quite eliminate -- lionfish abundances at local scales
\citep{deleon_2013,sandel_2015}. Complete eradication of lionfish
through fishing is unlikely because of their rapid recovery rates and
ongoing recruitment to shallow-water areas from persistent populations
in mesophotic ecosystems \citep{barbour_2011,andradibrown_2017}.
However, promoting lionfish consumption might create a level of demand
capable of incentivizing a stable fishery while controlling
shallow-water populations, thus creating alternative livelihoods and
avoiding further negative effects to local biota.

The feasibility of establishing fisheries through lionfish removal
programs has been extensively evaluated through field observations and
empirical modeling
\citep{barbour_2011,morris_2011,deleon_2013,johnston_2015,sandel_2015,usseglio_2017}.
Determining the feasibility of such initiatives requires modeling the
change in biomass in response to changes in fishing mortality
(\emph{i.e.} culling). A common way to model this is via
length-structured population models, where fish lengths are converted to
weight to calculate total biomass
\citep{barbour_2011,cote_2014,andradibrown_2017}. The allometric
length-weight relationship is thus an essential component of these
models, but this relationship can vary across regions as a response to
biotic and abiotic conditions \citep{johnson_2016}.

Outcomes of previous studies suggest lionfish are likely to exhibit
spatial heterogeneity in the length-weight relationship for two main
reasons. First, fish with sedentary behavior such as lionfish often
exhibit high levels of spatial variation in important life history
characteristics such as growth or natural mortality rates
\citep{gunderson_2008,hutchinson_2008,wilson_2012,guan_2013}. In fact,
culling programs are effective in reducing local adult populations
largely because lionfish have high levels of site fidelity and small
home ranges \citep{Fishelson_1997,kochzius_2005,jud_2012,cote_2014}.
Second, genetic analysis of lionfish suggests biological differences due
to the existence of two genetically distinct invasive subpopulations
between the Western Atlantic and the Caribbean \citep{betancurr_2011}.
Site-specific studies that calculate the length-weight relationship of
lionfish report variable estimates, and these differences may be
increasingly important when estimating the potential effectiveness of
lionfish culling programs
\citep{barbour_2011,morris_2011,cote_2014,johnston_2015}. However, the
influence of using \emph{ex situ} parameters when estimating the
length-weight relationship remains unexplored.

Our objective was to quantify the magnitude of error caused by using
\emph{ex situ} parameter values when estimating lionfish weight from
length observations. In this study, we calculated and reported the first
length-weight relationship for lionfish in the central Mexican Caribbean
(n = 109). We then estimated lionfish weight in this area using
previously published length-weight relationships for lionfish
populations in North Carolina, the Northern and Southern Gulf of Mexico,
the Southern Mexican Caribbean, the Bahamas, Little Cayman, Jamaica,
Bonaire, Puerto Rico, and Costa Rica. By comparing these weight
estimates to the \emph{in situ} length-weight observations, we showed
that using \emph{ex situ} parameter values resulted in up to a threefold
under- or overestimation of lionfish weight.

\section*{Methods}

We reviewed 12 published studies and obtained 17 length-weight
relationships for the Western Atlantic (n = 2), Gulf of Mexico (n = 7),
and Caribbean (n = 8, Table \ref{tab:all_params}, Fig \ref{fig:map}). We
collected information on sex differentiation, location, length and depth
ranges, and sampling methods from each study when available. Only two
studies reported parameters for each sex
\citep{aguilarperera_2016,fogg_2013}, so we assumed data from both sexes
were combined in a study if sex was unspecified. Reviewed studies
presented information for organisms ranging from 25-475 mm in Total
Length (\(TL\)) and were obtained at depths between 0.5 m and 57 m. Four
studies explicitly stated that their organisms were sampled with pole
spears \citep{dahl_2014,aguilarperera_2016,chin_2016,sabidoitz_2016},
and six studies mentioned that some of their organisms were obtained
with pole spears (or other type of harpoon) but also hand-held nets or
fish traps
\citep{barbour_2011,fogg_2013,edwards_2014,toledohernndez_2014,sandel_2015,sabidoitza_2016}.
Two studies did not specify how organisms were sampled
\citep{darling_2011,deleon_2013}.

We also used data from 109 lionfish sampled by
\citet{villaseorderbez_2014}, who used hand nets and numbered bottles to
collect organisms from 10 sampling sites along the central Mexican
Caribbean coast in the Summer of 2010 (Supplementary Table 1). Sampling
locations included wall and carpet reefs at depths between 5.7 m and
38.1 m. The use of hand nets prevented any weight loss due to bleeding
and allowed better representation of small sizes by avoiding gear
selectivity. Organisms were euthanized via pithing and Total Length (TL;
mm) and Total Weight (TW; g) were recorded.

\clearpage

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/map-1.pdf}
\caption{\label{fig:map}Locations where allometric growth parameters of
lionfish (\emph{Pterois spp}) have been reported. Circle sizes indicate
sample size from each study, colors indicate the \(b\) coefficient from
Eq. \ref{eq:allometric}.}
\end{figure}

The weight-at-length relationship for lionfish in the central Mexican
Caribbean was calculated with the allometric growth function

\begin{equation}
\label{eq:allometric}
TW = aTL^b
\end{equation}

where \(a\) is the ponderal index and \(b\) is the scaling exponent or
allometric parameter. We linearized the equation using
\(log_{10}\)-transformation and estimated the coefficients using an
Ordinary Least Squares Regression with a heteroskedastic-robust standard
error correction \citep{zeileis_2004}. Coefficients were tested with a
two-tailed Student's t-test, and the significance of the regression was
corroborated with an F-test.

Some of the reviewed studies inconsistently defined \(a\) as either the
ponderal index from Eq. \ref{eq:allometric} or the y-intercept from the
linearized log-transformed equation. Other studies incorrectly reported
parameters as mm-to-g conversions when they were in fact cm-to-g
conversions. We standardized each study by converting coefficients and
report all parameters as TL (mm) to TW (g) conversions. Locations where
allometric studies have been performed are shown in Figure \ref{fig:map}
and summarized in Table \ref{tab:all_params}.

We obtained a total of 18 parameter pairs by combining length-weight
parameters extracted from the literature and the additional pair
calculated here. Recall that the objective of this study is not to
describe variations between populations, but rather to estimate how the
use of \emph{ex situ} parameters influences the accuracy of weight
estimates for lionfish, using the central Mexican Caribbean as a case
study. Using each of the 18 parameter pairs, we estimated \(TW\) from
the \(TL\) observations collected in the central Mexican Caribbean (n =
109, with \(TL\) ranging from 34 mm to 310 mm) and divided predicted
weights by known observed weights to obtain a simple measure of over- or
underestimation. Difference in mean weight ratios were tested with an
analysis of covariance (ANCOVA), where the sources of variation were the
study and \(TL\). Ratios were logit-transformed prior to analysis, and a
\emph{post-hoc} Tukey's test was used to identify groups where mean
ratios did not differ. All analyses were performed in R version 3.5.2
\citep{rcore_2018}. Raw data and code used in this work are available on
github at github.com/jcvdav/lionfish\_biometry.

\begin{table}[t]

\caption{\label{tab:unnamed-chunk-2}\label{tab:all_params}Summary of 18 allometric growth parameters available for lionfish in the invaded range from peer-reviewed literature and this study. All parameters have been adjusted to convert from millimeters to grams. n = Sample size, Sex specifies whether data was presented for Females (F), Males (M), or both sexes combined (B), a = scaling parameter (presented in $\times 10^{-5}$), b = exponent.}
\centering
\begin{tabular}{llllrll}
\toprule
Region & Sex & n & a & b & $R^2$ & Reference\\
\midrule
Western Atlantic & B & 774 & 2.90 & 2.89 & - & Barbour et al., 2011\\
Western Atlantic & B & - & 0.25 & 3.29 & - & Darling et al., 2011\\
GoM & B & 934 & 0.21 & 3.34 & 0.98 & Dahl \& Patterson, 2014\\
GoM & B & 472 & 0.29 & 3.30 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & F & 67 & 0.12 & 3.47 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
\addlinespace
GoM & M & 59 & 0.42 & 3.23 & 0.95 & Aguilar-Perera \& Quijano-Puerto, 2016\\
GoM & B & 582 & 0.14 & 3.43 & 0.99 & Fogg et al., 2013\\
GoM & M & 119 & 0.27 & 3.31 & 0.97 & Fogg et al., 2013\\
GoM & F & 115 & 0.68 & 3.14 & 0.94 & Fogg et al., 2013\\
Caribbean & B & 458 & 3.60 & 2.81 & - & Sandel et al., 2015\\
\addlinespace
Caribbean & B & 419 & 2.80 & 2.85 & 0.87 & Chin et al., 2016\\
Caribbean & B & 1450 & 2.30 & 2.89 & 0.92 & de Leon et al., 2013\\
Caribbean & B & 1887 & 0.30 & 3.24 & 0.97 & Edwards et al., 2014\\
Caribbean & B & 2143 & 0.52 & 3.18 & 0.99 & Sabido-Itza et al., 2016\\
Caribbean & B & 227 & 0.80 & 3.11 & 0.96 & Toledo-Hernández et al., 2014\\
\addlinespace
Caribbean & B & 449 & 0.23 & 3.25 & 0.97 & Sabido-Itza et al., 2016b\\
Caribbean & B & 368 & 0.32 & 3.19 & 0.98 & Sabido-Itza et al., 2016b\\
Caribbean & B & 109 & 0.32 & 3.23 & 0.98 & This study\\
\bottomrule
\end{tabular}
\end{table}

\section*{Results}

The length-weight relationship for organisms from the central Mexican
Caribbean resulted in coefficient values of \(a = 3.206e-06\) and
\(b = 3.235\) (\(R^2 = 0.977\), \(F_{1, 107} = 6928.67; p < 0.001\)).
The allometric factor (\(b\)) was significantly different from \(b = 3\)
(\(t_{107} = 6.04; p<0.001\)) corroborating that lionfish present
allometric growth. The length-weight coefficients estimated here were
within the range identified by studies from other regions (Table
\ref{tab:all_params}). Figure \ref{fig:l-w-carib} shows the relationship
between TL and TW for this region.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/fit1-1.pdf}
\caption{\label{fig:l-w-carib}Length-weight relationship for 109
lionfish sampled in the central Mexican Caribbean. Points indicate
samples, dashed black line indicates curve of best fit, marginal plots
represent the density distribution of each variable.}
\end{figure}

There were significant differences in our predicted weights for the
central Mexican Caribbean when using each of the different pairs of
parameters (\(F_{17, 1943} = 24.96; p < 0.001\)). The lowest weight
estimates for the observed lengths resulted from using the allometric
parameters from Banco Chinchorro in the Caribbean, with mean \(\pm\) SD
of 40.37 \(\pm\) 58.74 g \citep{sabidoitz_2016}. In contrast, the
highest weight estimates came from the Western Atlantic with 73.76
\(\pm\) 96.11 g \citep{barbour_2011}. To put this in context, true
observed weights have a mean of 52.56 \(\pm\) 76.58 g. Weights predicted
from these extreme parameters correspond to predicted-to-observed weight
ratios of 0.80 \(\pm\) 0.19 and 1.76 \(\pm\) 0.50 (mean \(\pm\) SD),
respectively. The largest under- and overestimations resulted in ratios
of 0.36 and 3.51 of the actual observed weight, indicating that \emph{ex
situ} parameters can result in highly erroneous weight estimates.

Tukey's \emph{post-hoc} test suggests that weight ratios for the central
Mexican Caribbean were not different from those obtained with parameters
from Little Cayman, the Bahamas, and some sites in the Gulf of Mexico
(Tukey's HSD \(p > 0.05\)). All weight estimates using parameters from
the Gulf of Mexico and Western Atlantic were higher than observed
values, and only parameters from the Caribbean produced weights smaller
than observed (Fig \ref{fig:bio_ratio}). The regional average (\(\pm\)
SD) of predicted-to-observed weight ratios from these three regions were
1.24 \(\pm\) 0.309, 1.41 \(\pm\) 0.523, and 1.20 \(\pm\) 0.423 for the
Gulf of Mexico, Western Atlantic, and Caribbean, respectively. This
suggests that the smallest errors are observed when using parameters
from other locations in the Caribbean. Predicted-to-observed weight
ratios are presented in Figure \ref{fig:bio_ratio}.

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/fit2-1.pdf}
\caption{\label{fig:all_allo}Length-weight relationships (n = 18) for 12
studies and this study. The curves are shown for the range of lengths
reported in each study (See Supplementary Table 2); when ranges were not
present, we use the ones found in this study (34 mm - 310 mm). Colors
indicate studies from which the parameters were extracted. Dotted,
dashed, and solid lines show models for males, females, and combined
sexes, respectively. Letters in parentheses indicate if the study comes
from the Gulf of Mexico (GoM), Western Atlantic (WA), or Caribbean (Ca).
The dashed black line represents the relationship estimated in this
study. There are two solid green lines for Sabido-Itza et al, 2016b, one
for each of the two sites for which they report parameters. A log-log
version of this figure is presented in Figure S4.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/pred_obs-1.pdf}
\caption{\label{fig:bio_ratio}Violin plot of predicted-to-observed
weight ratios for 18 pairs of allometric parameters. Blue circles
indicate median values and like letters indicate values that do not
differ significantly. For Sabido-Itza et al, 2016b, BC and PNAX make
reference to Banco Chinchorro and Parque Nacional Arrecifes de Xcalak,
two sites for which they report parameters.}
\end{figure}

\clearpage

\section*{Discussion}

Our results suggest that lionfish exhibit highly variable, spatially
heterogeneous allometric relationships across their invaded range, and
that this variation is relevant for managing invasions. Moreover, we
show that the use of \emph{ex situ} parameter values may lead to highly
biased weight estimates. Our comparison of observed weights to those
predicted with locally-informed parameters and \emph{ex situ} parameters
showed that weight can be overestimated by more than threefold,
highlighting the need to use local information. Here we discuss the
implications of our findings, possible shortcommings in our analyses,
and highlight potential future research directions.

Differences in length-weight relationships have traditionally been
highlighted as potential pitfalls to fishery management. For example,
\citet{wilson_2012} show that small-scale variations in length-at-age
and fishing mortality in other Scorpaeniformes translate to differential
landings, effort, and catch per unit effort in the live fish fishery of
California, and that these differences must be taken into account in
management plans. The lionfish case poses the opposite scenario, where
the manager desires to eradicate the species. To accurately gauge both
the effectiveness of lionfish removal efforts and the resources needed
to successfully manage an invasion, we must acknowledge and understand
regional biological differences in important variables such as
allometric growth parameters.

We detected substantial differences in weight-at-length between
organisms from the Caribbean, Gulf of Mexico, and Western Atlantic.
Groupings of predicted-to-observed weight ratios identified in our
\emph{post hoc} testing aligned with the spatial distribution of the
examined studies, suggesting that these differences may be mediated by
space. These regional allometric differences mirror similar patterns in
length-at-age of lionfish across both their invaded and native regions
\citep{pusack_2016}. Variation may be driven by genetics or by
organisms' exposure to distinct environmental conditions. For example,
\citet{betancurr_2011} used mitochondrial DNA to demonstrate the
existence of two distinct population groups, identified as the
``Caribbean group'' and ``Northern Group'', and \citet{fogg_2015}
alternatively suggested that length-at-age differences may be driven by
the environment.

We might be inclined to attribute all variation in the lionfish
length-weight relationship to the spatial origin of these parameters.
However, samples from the 12 studies included here were not only
collected in different locations, but also at different points in time
and across different depth and size ranges (See Supplementary Table 2
for an extended version of Table 1 that incorporates these variables
when available). While we could not evaluate how these factors influence
previous estimates without the necessary raw data from every study, we
still show that the lack of locally-calculated parameters can induce
significant bias when calculating weight from length observations. The
magnitude of the bias discovered in this study and our lack of
understanding the sources driving spatial variation for lionfish
highlights the need to simultaneously collect length-weight information
across the invaded range to test for spatially-induced patterns and to
link these findings to previously suggested environmental and genetic
structures. Such an endeavor would provide insight into lionfish biology
and better inform management.

Applying parameters estimates to lengths outside the range of lengths
originally used to estimate the parameters may also induce error. Our
smallest observed organism was 34 mm in \(TL\), and only two studies
estimated parametrs with smaller organisms
\citep{edwards_2014,sabidoitza_2016}. However, our largest organism had
a \(TL\) of 310 mm, which is well within the range of all other studies
(the next smallest maximum length was 325 mm; See Supplementary Table
2). Due to the power function describing the allometric relationship
(\emph{i.e.} Eq. 1), the error is larger when extrapolation is done for
lengths that are larger than the maximum length used to estimate the
parameters.

The results presented here have key implications for management. For
example, \citet{edwards_2014} simulated a lionfish culling program under
two scenarios, one using length-at-age and length-to-weight parameters
from North Carolina and one using parameters from Little Cayman. Their
results show that using different parameters caused up to a four-year
difference in the time required for the simulated lionfish population to
recover to 90\% of its initial biomass after removals ceased. Here, we
show that using one set of length-weight parameters versus another for a
given length can result in more than a threefold under- or
overestimation of total weight. These differences become especially
important when allocating resources for lionfish removal programs,
incentivizing lionfish fisheries as a source of alternative livelihoods,
or estimating ecosystem impacts. Research efforts focused on invasive
lionfish populations need to use parameters calculated for their region
to the extent possible, or use different sets of parameters that provide
appropriate upper and lower bounds in their results.

\section*{Acknowledgements}

We thank Nils Van Der Haar and Michael Doodey from Dive Aventuras as
well as Guillermo Lotz-Cador who provided help to collect samples. We
are grateful for comments raised by the editor and two anonymous
reviewers, which significantly increased the quality of this work.

Conflict of Interest: The authors declare that they have no conflict of
interest.

\bibliography{references}

\end{document}